---
title: "Scrape.it for Final Project"
author: "Sam Tolbert"
date: "2024-11-27"
output:
  pdf_document: default
  html_document: default
---


testing testing git

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rvest)
library(purrr)
library(lubridate)
library(here)
library(dplyr)

#install.packages("dataRetrieval")
library(dataRetrieval)

install.packages("forecast")

```

Retrieving data for Sugar Grove Gage from USGS using USGS's dataRetrieval package.
I find its siteNumber is 03479000 from "https://waterdata.usgs.gov/monitoring-location/03479000/#parameterCode=
00060&period=P7D&showMedian=false".
```{r echo=TRUE}

siteNumberSugarGrove<-('03479000')

#I find the MetaData for the site
SugarGroveMeta <- whatNWISsites(siteNumbers = siteNumberSugarGrove)
print(SugarGroveMeta) 

#I then use sitedata function to look at what data is available
SugarGroveSiteData<-suppressWarnings(whatNWISdata(siteNumber=siteNumberSugarGrove))
```

Looking at what data is available, I see that the parameter codes (parmcodes)
for stream discharge (00060) , precipitation (00045), and gage height (0065).
I also see that our relevant stat codes are daily mean (0003) and sum (00006)
#

```{r message=FALSE, warning=FALSE}
pcodes<-c('00060','00045','00065') #parameters discharge, precip, and gage height
scode<-c('00003', '00006') #daily mean and daily accumulated


SugarGroveData<-readNWISdata( #dataRetrival function from USGS API
  siteNumbers = siteNumberSugarGrove,
  parameterCd= pcodes,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)
  
#View(SugarGroveData)  
```

This is my data! However, I need to clean this up. 
First I want to remove the 'cd' columns that indicate if the data is official or 
provisional as this won't have a significant effect on our findings. 
We can mention that in our discussion of the meta data but its not 
worth a column in our df. 

Then I want to clearly label each column with its parameter and stat code:

```{r}
CleanedDataSugarGrove<-SugarGroveData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )

```

I want to repeat this data retrieval and cleaning for two additional sites,
French Broad River in Asheville in Southwestern NC (03451500) 
and East Fork Pigeon River in far Southwestern NC near Canton (03456500) for
examinations of 3 different rivers in different parts of the state.

First I check the metadata and sitedata to see if similar data and parameters
are availale to compare to Sugar Grove

```{r}

#First the Asheville site

siteNumberAsheville<-('03451500')

#I find the MetaData for the site
AshevilleMeta <- whatNWISsites(siteNumbers = siteNumberAsheville)
print(AshevilleMeta) 

#I then use sitedata function to look at what data is available
AshevilleSiteData<-whatNWISdata(siteNumber=siteNumberAsheville)

```

Then I decide I want to create an easy data retrieval function for my chosen
parameters based on my previous retreival.

```{r}

WaterData.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
df_name<-paste0('CleanData_', siteNo)
assign(df_name, CleanData, envir= .GlobalEnv)

return(CleanData)
}
```

I run this function for Asheville's siteNo
```{r}
siteNoAsheville<-'03451500'

siteNo<-siteNoAsheville
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataAsheville<-WaterData.Function(siteNo, pcode, scode)
#View(CleanedDataAsheville)


```

I then do the same for the Pigeon River Site and its site number

```{r}

siteNoPigeonRiver<-'03456500'

siteNo<-siteNoPigeonRiver
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataPigeonRiver<-WaterData.Function(siteNo, pcode, scode)
#View(CleanedDataPigeonRiver)

```

```{r DECOMPOSITION PLOTS: Chrissie-ASHEVILLE}

library(dplyr)
library(lubridate)
library(zoo)

# Convert dateTime to Date class
CleanedDataAsheville <- CleanedDataAsheville %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values
CleanedDataAsheville.clean <- CleanedDataAsheville %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Generate time series objects
CleanedDataAsheville.monthly <- CleanedDataAsheville.clean %>%
  filter(!is.na(mean_discharge))

# Step 1: Add columns for year and month, then calculate monthly mean discharge
CleanedDataAsheville.monthly <- CleanedDataAsheville.clean %>%
  mutate(
    year = year(dateTime),   # Extract year from dateTime
    month = month(dateTime)  # Extract month from dateTime
  ) %>%
  group_by(year, month) %>%
  summarize(
    mean_discharge = mean(Mean.Discharge.clean, na.rm = TRUE)
  ) %>%
  ungroup()

# Step 2: Create a new Date column for graphing, setting each month-year combination to the first day of the month
CleanedDataAsheville.monthly <- CleanedDataAsheville.monthly %>%
  mutate(
    dateTime = as.Date(paste(year, month, "01", sep = "-")))

# Extract start and end years and months
s_year <- year(first(CleanedDataAsheville$dateTime))
f_year <- year(last(CleanedDataAsheville$dateTime))
s_month <- month(first(CleanedDataAsheville$dateTime))
f_month <- month(last(CleanedDataAsheville$dateTime))

# 1. Generate the daily time series object
CleanedDataAsheville.daily.ts <- ts(
  CleanedDataAsheville.clean$Mean.Discharge.clean,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# 2. Generate the monthly time series object using aggregated data
CleanedDataAsheville.monthly.ts <- ts(
  CleanedDataAsheville.monthly$mean_discharge,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 12
)

# Display the created time series
CleanedDataAsheville.daily.ts
CleanedDataAsheville.monthly.ts

# Decompose the daily time series
CleanedDataAsheville.daily.decomp <- stl(CleanedDataAsheville.daily.ts, s.window = "periodic")

# Plot the components of the daily decomposition
plot(CleanedDataAsheville.daily.decomp, main = "Decomposition of Daily Discharge for Asheville")

# Decompose the monthly time series
CleanedDataAsheville.monthly.decomp <- stl(CleanedDataAsheville.monthly.ts, s.window = "periodic")

# Plot the components of the monthly decomposition
plot(CleanedDataAsheville.monthly.decomp, main = "Decomposition of Monthly Discharge for Asheville")


```

```{r DECOMPOSITION PLOTS: Chrissie-PIGEON RIVER}

library(dplyr)
library(lubridate)
library(zoo)

# Convert dateTime to Date class
CleanedDataPigeonRiver <- CleanedDataPigeonRiver %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Generate time series objects
CleanedDataPigeonRiver.monthly <- CleanedDataPigeonRiver.clean %>%
  filter(!is.na(mean_discharge))

# Step 1: Add columns for year and month, then calculate monthly mean discharge
CleanedDataPigeonRiver.monthly <- CleanedDataPigeonRiver.clean %>%
  mutate(
    year = year(dateTime),   # Extract year from dateTime
    month = month(dateTime)  # Extract month from dateTime
  ) %>%
  group_by(year, month) %>%
  summarize(
    mean_discharge = mean(Mean.Discharge.clean, na.rm = TRUE)
  ) %>%
  ungroup()

# Step 2: Create a new Date column for graphing, setting each month-year combination to the first day of the month
CleanedDataPigeonRiver.monthly <- CleanedDataPigeonRiver.monthly %>%
  mutate(
    dateTime = as.Date(paste(year, month, "01", sep = "-")))

# Extract start and end years and months
s_year <- year(first(CleanedDataPigeonRiver$dateTime))
f_year <- year(last(CleanedDataPigeonRiver$dateTime))
s_month <- month(first(CleanedDataPigeonRiver$dateTime))
f_month <- month(last(CleanedDataPigeonRiver$dateTime))

# 1. Generate the daily time series object
CleanedDataPigeonRiver.daily.ts <- ts(
  CleanedDataPigeonRiver.clean$Mean.Discharge.clean,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# 2. Generate the monthly time series object using aggregated data
CleanedDataPigeonRiver.monthly.ts <- ts(
  CleanedDataPigeonRiver.monthly$mean_discharge,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 12
)

# Display the created time series
CleanedDataPigeonRiver.daily.ts
CleanedDataPigeonRiver.monthly.ts

# Decompose the daily time series
CleanedDataPigeonRiver.daily.decomp <- stl(CleanedDataPigeonRiver.daily.ts, s.window = "periodic")

# Plot the components of the daily decomposition
plot(CleanedDataPigeonRiver.daily.decomp, main = "Decomposition of Daily Discharge for Pigeon River")

# Decompose the monthly time series
CleanedDataPigeonRiver.monthly.decomp <- stl(CleanedDataPigeonRiver.monthly.ts, s.window = "periodic")

# Plot the components of the monthly decomposition
plot(CleanedDataPigeonRiver.monthly.decomp, main = "Decomposition of Monthly Discharge for Pigeon River")

```

```{r DECOMPOSITION PLOTS: Chrissie-SUGAR GROVE}

library(dplyr)
library(lubridate)
library(zoo)

# Convert dateTime to Date class
CleanedDataSugarGrove <- CleanedDataSugarGrove %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Generate time series objects
CleanedDataSugarGrove.monthly <- CleanedDataSugarGrove.clean %>%
  filter(!is.na(mean_discharge))

# Step 1: Add columns for year and month, then calculate monthly mean discharge
CleanedDataSugarGrove.monthly <- CleanedDataSugarGrove.clean %>%
  mutate(
    year = year(dateTime),   # Extract year from dateTime
    month = month(dateTime)  # Extract month from dateTime
  ) %>%
  group_by(year, month) %>%
  summarize(
    mean_discharge = mean(Mean.Discharge.clean, na.rm = TRUE)
  ) %>%
  ungroup()

# Step 2: Create a new Date column for graphing, setting each month-year combination to the first day of the month
CleanedDataSugarGrove.monthly <- CleanedDataSugarGrove.monthly %>%
  mutate(
    dateTime = as.Date(paste(year, month, "01", sep = "-")))

# Extract start and end years and months
s_year <- year(first(CleanedDataSugarGrove$dateTime))
f_year <- year(last(CleanedDataSugarGrove$dateTime))
s_month <- month(first(CleanedDataSugarGrove$dateTime))
f_month <- month(last(CleanedDataSugarGrove$dateTime))

# 1. Generate the daily time series object
CleanedDataSugarGrove.daily.ts <- ts(
  CleanedDataSugarGrove.clean$Mean.Discharge.clean,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# 2. Generate the monthly time series object using aggregated data
CleanedDataSugarGrove.monthly.ts <- ts(
  CleanedDataSugarGrove.monthly$mean_discharge,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 12
)

# Display the created time series
CleanedDataSugarGrove.daily.ts
CleanedDataSugarGrove.monthly.ts

# Decompose the daily time series
CleanedDataSugarGrove.daily.decomp <- stl(CleanedDataSugarGrove.daily.ts, s.window = "periodic")

# Plot the components of the daily decomposition
plot(CleanedDataSugarGrove.daily.decomp, main = "Decomposition of Daily Discharge for Sugar Grove")

# Decompose the monthly time series
CleanedDataSugarGrove.monthly.decomp <- stl(CleanedDataSugarGrove.monthly.ts, s.window = "periodic")

# Plot the components of the monthly decomposition
plot(CleanedDataSugarGrove.monthly.decomp, main = "Decomposition of Monthly Discharge for Sugar Grove")

```
7-day average


```{r ASHEVILLE}
library(dplyr)
library(lubridate)
library(zoo)


# Convert dateTime to Date class and clean columns
CleanedDataAsheville <- CleanedDataAsheville %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataAsheville.clean <- CleanedDataAsheville %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataAsheville.clean <- CleanedDataAsheville.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataAsheville.clean <- CleanedDataAsheville.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataAsheville.clean$dateTime))
s_month <- month(first(CleanedDataAsheville.clean$dateTime))
f_year <- year(last(CleanedDataAsheville.clean$dateTime))
f_month <- month(last(CleanedDataAsheville.clean$dateTime))

# Re-create the daily time series object
CleanedDataAsheville.daily.7DayAvg.ts <- ts(
  CleanedDataSugarGrove.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataAsheville.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataAsheville.daily.7DayAvg.decomp <- stl(
  CleanedDataAsheville.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataAsheville.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Asheville"
)


```

```{r SUGAR GROVE}
library(dplyr)
library(lubridate)
library(zoo)


# Convert dateTime to Date class and clean columns
CleanedDataSugarGrove <- CleanedDataSugarGrove %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataSugarGrove.clean$dateTime))
s_month <- month(first(CleanedDataSugarGrove.clean$dateTime))
f_year <- year(last(CleanedDataSugarGrove.clean$dateTime))
f_month <- month(last(CleanedDataSugarGrove.clean$dateTime))

# Re-create the daily time series object
CleanedDataSugarGrove.daily.7DayAvg.ts <- ts(
  CleanedDataSugarGrove.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataSugarGrove.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataSugarGrove.daily.7DayAvg.decomp <- stl(
  CleanedDataSugarGrove.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataSugarGrove.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Sugar Grove"
)


```
```{r PIGEON RIVER}
library(dplyr)
library(lubridate)
library(zoo)


# Convert dateTime to Date class and clean columns
CleanedDataPigeonRiver <- CleanedDataPigeonRiver %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataPigeonRiver.clean$dateTime))
s_month <- month(first(CleanedDataPigeonRiver.clean$dateTime))
f_year <- year(last(CleanedDataPigeonRiver.clean$dateTime))
f_month <- month(last(CleanedDataPigeonRiver.clean$dateTime))

# Re-create the daily time series object
CleanedDataPigeonRiver.daily.7DayAvg.ts <- ts(
  CleanedDataPigeonRiver.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataPigeonRiver.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataPigeonRiver.daily.7DayAvg.decomp <- stl(
  CleanedDataPigeonRiver.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataPigeonRiver.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Pigeon River"
)


```



