---
title: "Final Project Draft"
author: "Sam Tolbert, Chrissie Pantoja, Ana Andino, Gretchen Barbera"
date: "2024-12-09"
output:
  html_document:
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rvest)
library(purrr)
library(lubridate)
library(here)
library(dplyr)
library(zoo)
library(ggplot2)
library(tidyr)

#install.packages("dataRetrieval")
library(dataRetrieval)

#install.packages("forecast")
library(forecast)

```


Retrieving data for Sugar Grove Gage from USGS using USGS's "dataRetrieval" package.
I find its siteNumber is 03479000 from "https://waterdata.usgs.gov/monitoring-location/03479000/#parameterCode=
00060&period=P7D&showMedian=false".


<span style="font-size:15px; font-weight:bold;">Retrieving Data</span>


**Sugar Grove**
```{r echo=TRUE, warning=FALSE}

siteNumberSugarGrove<-('03479000')

#I find the MetaData for the site
SugarGroveMeta <- whatNWISsites(siteNumbers = siteNumberSugarGrove)
print(SugarGroveMeta) 

#I then use sitedata function to look at what data is available
SugarGroveSiteData<-suppressWarnings(whatNWISdata(siteNumber=siteNumberSugarGrove))
```

Looking at what data is available, I see that the parameter codes (parmcodes)
for stream discharge (00060) , precipitation (00045), and gage height (0065).
I also see that our relevant stat codes are daily mean (0003) and sum (00006)
#

```{r message=FALSE, warning=FALSE}
pcodes<-c('00060','00045','00065') #parameters discharge, precip, and gage height
scode<-c('00003', '00006') #daily mean and daily accumulated


SugarGroveData<-readNWISdata( #dataRetrival function from USGS API
  siteNumbers = siteNumberSugarGrove,
  parameterCd= pcodes,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)
  
#View(SugarGroveData)  
```

This is my data! However, I need to clean this up. 
First I want to remove the 'cd' columns that indicate if the data is official or 
provisional as this won't have a significant effect on our findings. 
We can mention that in our discussion of the meta data but its not 
worth a column in our df. 

Then I want to clearly label each column with its parameter and stat code:

```{r}
CleanedDataSugarGrove<-SugarGroveData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
#View(CleanedDataSugarGrove)


```

I want to repeat this data retrieval and cleaning for two additional sites,
French Broad River in Asheville in Southwestern NC (03451500) 
and East Fork Pigeon River in far Southwestern NC near Canton (03456500) for
examinations of 3 different rivers in different parts of the state.

First I check the metadata and sitedata to see if similar data and parameters
are availale to compare to Sugar Grove

**Asheville**

```{r, warning=FALSE}

#First the Asheville site

siteNumberAsheville<-('03451500')

#I find the MetaData for the site
AshevilleMeta <- whatNWISsites(siteNumbers = siteNumberAsheville)
print(AshevilleMeta) 

#I then use sitedata function to look at what data is available
AshevilleSiteData<-whatNWISdata(siteNumber=siteNumberAsheville)

```

Then I decide I want to create an easy data retrieval function for my chosen
parameters based on my previous retreival.

```{r}

WaterData.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
df_name<-paste0('CleanData_', siteNo)
assign(df_name, CleanData, envir= .GlobalEnv)

return(CleanData)
}
```

I run this function for Asheville's siteNo
```{r}
siteNoAsheville<-'03451500'

siteNo<-siteNoAsheville
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataAsheville<-WaterData.Function(siteNo, pcode, scode)
#View(CleanedDataAsheville)


```

I then do the same for the Pigeon River Site and its site number
**Pigeon River**

```{r}

siteNoPigeonRiver<-'03456500'

siteNo<-siteNoPigeonRiver
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataPigeonRiver<-WaterData.Function(siteNo, pcode, scode)

#View(CleanedDataPigeonRiver)
```

<span style="font-size:15px; font-weight:bold;">Research Question</span>

<span style="font-size:20px; font-weight:bold;">How unusual was Helene?</span>


First let's create separate layers from our data for 2024 and
then Helene as an event within 2024 a 7 day span from 9/22/24 to 9/29/24.
**Investigating Helene**

```{r Helene Precipitation event over 2024}

#first we create our df for 2024
Data2024_sg <- CleanedDataSugarGrove %>%
  filter(year(dateTime) == 2024)

Data2024_ash <- CleanedDataAsheville %>%
  filter(year(dateTime) == 2024)

Data2024_pg <- CleanedDataPigeonRiver %>%
  filter(year(dateTime) == 2024)




#then we isolate the Helene as a 7 day event:

HeleneData_sg <- CleanedDataSugarGrove %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))

HeleneData_ash <- CleanedDataAsheville %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))

HeleneData_pg <- CleanedDataPigeonRiver %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))


#View(HeleneData_pg)


```


Then, let's see how the week of Helene compares to the rest of 2024 within our
parameters

**Precipitation**
```{r, warning=FALSE}

# First: Sugar Grove Precipitation for 2024
ggplot(Data2024_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black", size = 2) +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black", size = 2) +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()

# Second: Asheville Precipitation for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Asheville",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()

# Third: Pigeon River Precipitation for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()




```



**Then discharge**
```{r Helene Discharge event over 2024, warning=FALSE}

# First: Sugar Grove Discharge for 2024
ggplot(Data2024_sg, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()


# Second: Asheville Discharge for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Asheville",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()




# Third: Pigeon River Discharge for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()


```



**Then Gage Height**
```{r}

# First: Sugar Grove Gage Height for 2024

ggplot(Data2024_sg, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()

# Second: Asheville Discharge for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Asheville",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()


# Third: Pigeon River Gage Height for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()




```


Okay, so clearly Helen was the maximum event for these parameters for 2024.


<span style="font-size:15px;font-style:italic;">How do these maximums compare to all historical data we have access to?</span>


Let's zoom out on the same dates but compared with all historical data.

**Precipitation**

```{r, warning=FALSE}

#First: Sugar Grive Historical Precipitation 

SugarGroveDailyPrecipitationLinePlot<-ggplot(CleanedDataSugarGrove%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x= dateTime,
   y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "lightblue", size = 1) +
  geom_point(data = HeleneData_sg, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +  
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Sugar Grove(1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 
 
print(SugarGroveDailyPrecipitationLinePlot) 


#Second: Asheville Historical Precipitation 

AshevilleDailyPrecipitationLinePlot<-ggplot(CleanedDataAsheville%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x = dateTime, 
  y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +    
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 

print(AshevilleDailyPrecipitationLinePlot)


#Third: Pigeon River Historical Precipitation 
PigeonRiverDailyTotalPrecipitationPlot<-ggplot(CleanedDataPigeonRiver%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x= dateTime,
   y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "grey", size = 1) +
  geom_point(data = HeleneData_pg, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +    
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Pigeon River (1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 

print(PigeonRiverDailyTotalPrecipitationPlot)



```


**Discharge**
```{r, warning=FALSE}

#First:Sugar Grove Historical Discharge Data

SugarGroveDailyDischargeLinePlot<-ggplot(CleanedDataSugarGrove%>%
    drop_na(`Mean Discharge`), aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="lightblue")+
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Sugar Grove(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()
 
print(SugarGroveDailyDischargeLinePlot) 



#Second: Asheville Historical Discharge Data 

AshevilleDailyDischargeLinePlot<-ggplot(CleanedDataAsheville %>%
    drop_na(`Mean Discharge`), aes(
  x = dateTime, 
  y = `Mean Discharge`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Mean Discharge (cubic meters per second)"
  ) +
  theme_minimal() 

print(AshevilleDailyDischargeLinePlot)


#Third: Pigeon River Historical Discharge Data 

PigeonRiverDailyDischargeLinePlot<-ggplot(CleanedDataPigeonRiver%>%
    drop_na(`Mean Discharge`), aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="grey")+
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Pigeon River(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()

print(PigeonRiverDailyDischargeLinePlot)


```





**Gage Height**
```{r, warning=FALSE}

#First: Sugar Grove Historical Gage Height 

SugarGroveDailyGageHeightPlot<-ggplot(CleanedDataSugarGrove %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "lightblue", size = 1) +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +    
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Sugar Grove Gage Height",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(SugarGroveDailyGageHeightPlot)


#Second: Asheville Historical Gage Height 

AshevilleDailyGageHeightPlot<-ggplot(CleanedDataAsheville %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +      
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(AshevilleDailyGageHeightPlot)


#Third: Pigeon River Historical Gage Height

PigeonRiverDailyGageHeightPlot<-ggplot(CleanedDataPigeonRiver %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "grey", size = 1) +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +      
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Pigeon River Gage Height",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(PigeonRiverDailyGageHeightPlot)


```

These graphs reaveal that Helene was responsible for some
outlier behavior, but it doesn't account for seasonality and other
trends.

Let's decompose to account for these trends

###CHRISSIE PUT YOUR DECOMPS HERE###












These graphs are interesting, and some of them show 2024 as an outlierbut they
don't say by exactly how much Helen was an outlier.

We want to be able to say with statistical certainty how unusual Helene was as
an event.

We will use the Hazen Method to find the recurrence intervals for Helene within
these parameters. We will do this below.


First we find the recurrence intervals of 7 days precipitation events. We know from
our first graphs that assuming Helen is our 7 day averaged maximum for 2024
is a reasonable assumption:

<span style="font-size:18px; font-weight:bold;">Helene Recurrence Intervals</span>

**Sugar Grove Hazen Precipitation**

```{r Sugar Grove Precip Hazen, warning=FALSE}
SugarGrovePrecip <- CleanedDataSugarGrove %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

SugarGrovePrecip <- SugarGrovePrecip %>%
  mutate(Year = year(dateTime))

SugarGrovePrecip <- SugarGrovePrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))

SugarGrovePrecip <- SugarGrovePrecip %>% 
  filter(MaxEvent > 0)

SugarGrovePrecip <- SugarGrovePrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

SugarGrovePrecipHazen2024 <- SugarGrovePrecip %>% filter(Year == 2024)
print(SugarGrovePrecipHazen2024)


```


**Asheville Hazen Precipitation**
```{r Asheville Precipitation Hazen, warning=FALSE}
AshevillePrecip <- CleanedDataAsheville %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

AshevillePrecip <- AshevillePrecip %>%
  mutate(Year = year(dateTime))

AshevillePrecip <- AshevillePrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm = TRUE))

AshevillePrecip <- AshevillePrecip %>% 
  filter(MaxEvent > 0)

AshevillePrecip <- AshevillePrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100 * (2 * rank - 1)) / (2 * n()))) %>% 
  mutate(hazen = 100 / Fa_Value)

AshevillePrecipHazen2024 <- AshevillePrecip %>% filter(Year == 2024)
print(AshevillePrecipHazen2024)


```

The Hazen number tells us that the 7 day average discharge observed in
Asheville in 2024 was a 54 year event, in other words it can be expected to
happen once every 54 years, or about a 2% chance of happening any one year. That
gives us a statistical number to match to our observed outliers. Let's find the 
recurrence intervals for 2024 for our other sites

**Pigeon River Hazen Precipitation**

```{r Pigeon River Precip Hazen, warning=FALSE}
PigeonRiverPrecip <- CleanedDataPigeonRiver %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

PigeonRiverPrecip <- PigeonRiverPrecip %>%
  mutate(Year = year(dateTime))

PigeonRiverPrecip <- PigeonRiverPrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm = TRUE))

PigeonRiverPrecip <- PigeonRiverPrecip  %>% 
  filter(MaxEvent > 0)

PigeonRiverPrecip <- PigeonRiverPrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100 * (2 * rank - 1)) / (2 * n()))) %>% 
  mutate(hazen = 100 / Fa_Value)

PigeonRiverPrecipHazen2024 <- PigeonRiverPrecip %>% filter(Year == 2024)

print(PigeonRiverPrecipHazen2024)
```

In Pigeon River, it Helene was a 52 year event in terms of total precipitation,
but in Sugar Grove only an 18 year event, meaning Helene's
precipitation was less unusual in Sugar Grove than in Asheville or Pigeon River.


Let's now automate this function for our other two parameters and run our sites.

**Automated Discharge**

```{r}

Hazen2024Discharge.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
CleanDataFlood<-CleanData%>% 
  mutate(moving_avg = rollapply(`Mean Discharge`, width = 7, FUN = mean, align = "center", fill = NA)) 

CleanDataFlood<- CleanDataFlood%>%
  mutate(Year = year(dateTime))

CleanDataFlood <- CleanDataFlood%>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))

CleanDataFlood <- CleanDataFlood %>% 
  filter(MaxEvent > 0)

CleanDataFlood <- CleanDataFlood %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

CleanDataFlood2024 <- CleanDataFlood %>% filter(Year == 2024)


return(CleanDataFlood2024)
}

```

*And now we run our sites through our discharge function*

```{r}

SugarGroveDischargeHazen2024 <- Hazen2024Discharge.Function(siteNumberSugarGrove, pcode, scode)

AshevilleDischargeHazen2024 <- Hazen2024Discharge.Function(siteNumberAsheville, pcode, scode)

PigeonRiverDischargeHazen2024 <- Hazen2024Discharge.Function(siteNoPigeonRiver, pcode, scode)

SugarGroveDischargeHazen2024
AshevilleDischargeHazen2024
PigeonRiverDischargeHazen2024



```

*Now our function for Gage Height*

```{r}
Hazen2024GageHeight.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
CleanDataGageHeight<-CleanData%>% 
  mutate(moving_avg = rollapply(`Mean Gage Height`, width = 7, FUN = mean, align = "center", fill = NA)) 

CleanDataGageHeight<- CleanDataGageHeight%>%
  mutate(Year = year(dateTime))

CleanDataGageHeight <- suppressWarnings(CleanDataGageHeight%>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))
)

CleanDataGageHeight <- CleanDataGageHeight %>% 
  filter(MaxEvent > 0)

CleanDataGageHeight <- CleanDataGageHeight %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

CleanDataGageHeight2024 <- CleanDataGageHeight %>% filter(Year == 2024)


return(CleanDataGageHeight2024)
}

#testing function

SugarGroveGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNumberSugarGrove, pcode, scode)

AshevilleGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNumberAsheville, pcode, scode)

PigeonRiverGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNoPigeonRiver, pcode, scode)

SugarGroveGageHeightHazen2024
AshevilleGageHeightHazen2024
PigeonRiverGageHeightHazen2024

```



**Map of Gage Sites**

I would like to be able to now see geographically how unusual Helene was for each
site:
```{r}



library(RColorBrewer)

# Create Data Frame with Hazen Values
hazen_data <- data.frame(
  Location = c("Pigeon River", "Asheville", "Sugar Grove"),
  Latitude = c(35.554, 35.595, 36.267),
  Longitude = c(-82.982, -82.551, -81.785),
  Precipitation_Hazen = sapply(list(PigeonRiverPrecipHazen2024$hazen, 
                                    AshevillePrecipHazen2024$hazen, 
                                    SugarGrovePrecipHazen2024$hazen), function(x) as.numeric(unlist(x))),
  Discharge_Hazen = sapply(list(PigeonRiverDischargeHazen2024$hazen, 
                                 AshevilleDischargeHazen2024$hazen, 
                                 SugarGroveDischargeHazen2024$hazen), function(x) as.numeric(unlist(x))),
  GageHeight_Hazen = sapply(list(PigeonRiverGageHeightHazen2024$hazen, 
                                  AshevilleGageHeightHazen2024$hazen, 
                                  SugarGroveGageHeightHazen2024$hazen), function(x) as.numeric(unlist(x)))
)





color_pal <- colorNumeric(
  palette = c("blue", "green", "red"),  # Custom palette: low -> medium -> high
  domain = range(
    hazen_data$Precipitation_Hazen,
    hazen_data$Discharge_Hazen,
    hazen_data$GageHeight_Hazen,
    na.rm = TRUE
  ),
  reverse = FALSE
)


leaflet(hazen_data) %>%
  addTiles() %>%
  
  # Precipitation Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = ~color_pal(Precipitation_Hazen),
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Precipitation Hazen: ", round(Precipitation_Hazen, 2)
    ),
    radius = 10, group = "Precipitation"
  ) %>%
  
  # Discharge Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = ~color_pal(Discharge_Hazen),
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Discharge Hazen: ", round(Discharge_Hazen, 2)
    ),
    radius = 10, group = "Discharge"
  ) %>%
  
  # Gage Height Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = ~color_pal(GageHeight_Hazen),
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Gage Height Hazen: ", round(GageHeight_Hazen, 2)
    ),
    radius = 10, group = "Gage Height"
  ) %>%
  
  # Add Layer Controls
  addLayersControl(
    overlayGroups = c("Precipitation", "Discharge", "Gage Height"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  
  # Add a Legend
  addLegend(
    "bottomright",
    pal = color_pal,
    values = range(
      hazen_data$Precipitation_Hazen,
      hazen_data$Discharge_Hazen,
      hazen_data$GageHeight_Hazen,
      na.rm = TRUE
    ),
    title = "Hazen Value (Recurrence Interval)",
    opacity = 0.7
  )


```

This is cool! I want to add a lot more sites now that I have this functioning.
I go and download more sites in Western NC, by randomly selecting one per county
for every counry in Western NC and find the Hazen numbers for each parameter
```{r}




```



I then add them to the heatmap
```{r}

#install.packages("leaflet.extras")
library(leaflet.extras)

leaflet(hazen_data) %>%
  addTiles() %>%
  addHeatmap(
    lng = ~Longitude,
    lat = ~Latitude,
    intensity = ~Precipitation_Hazen,
    blur = 20,
    max = max(hazen_data$Precipitation_Hazen, na.rm = TRUE),
    radius = 15
  ) %>%
  addLegend(
    "bottomright",
    pal = color_pal,
    values = hazen_data$Precipitation_Hazen,
    title = "Precipitation Hazen",
    opacity = 0.7
  )


leaflet(hazen_data) %>%
  addTiles() %>%
  addCircleMarkers(
    ~Longitude, ~Latitude,
    fillColor = ~color_pal(Discharge_Hazen),
    fillOpacity = 0.8,
    stroke = FALSE,
    radius = 10,
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Discharge Hazen: ", round(Discharge_Hazen, 2)
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = color_pal,
    values = hazen_data$Discharge_Hazen,
    title = "Discharge Hazen",
    opacity = 0.7
  )



```