---
title: "Title: Rainfall Trends Following the Impact of Hurricane Helene 
in Sugar Grove, Asheville, and Pigeon River"
author: "Sam Tolbert, Chrissie Pantoja, Ana Andino, Gretchen Barbera"
date: "2024-12-09"
output:
  html_document:
    toc: true
    toc_float: true
---
**Context and Motivation**
Hurricane Helene struck North Carolina in late September, rapidly intensifying
to a Category 4 storm with sustained winds of 140 mph. The hurricane brought 
catastrophic rainfall, widespread flooding, landslides, destructive winds, and 
tornados, causing severe disruptions to both the economy and daily life. Months
later, the state is still recovering from the storm’s far-reaching effects.
Particularly striking is that Helene followed closely on the heels of 
Hurricane Milton, a Category 5 storm that had already devastated the west coast
of the United States. These concurrent events underscore the increasing severity 
and frequency of hurricanes, highlighting the urgent need to better understand 
the mechanisms behind and the impacts of such extreme weather events.

The devastation caused by Hurricane Helene has left a profound and lasting 
impact on North Carolina. The storm wreaked unprecedented destruction, 
leaving thousands of homes either destroyed or severely damaged. Millions of 
residents were cut off from critical infrastructure, including water, 
electricity, telecommunications, and healthcare services (OSBM, 2024).
Key infrastructure, such as roadways and bridges, was also heavily damaged. 
Economically, the storm's effects will be felt for years to come. 
Preliminary estimates of the damage and recovery needs across the state total approximately $53.6 billion, which includes $41.1 billion in direct damages, $7.6 billion in indirect impacts, and $4.8 billion needed for strengthening and mitigation efforts (OSBM, 2024). These figures highlight the enormity of the storm’s impact compared to other recent hurricanes, such as Hurricane Florence in 2018, which resulted in $16.7 billion in damages. Beyond the immediate economic losses, the storm significantly altered the state’s physical landscape, eroding valuable topsoil, changing river courses, and submerging homes that had stood for generations. The long-term effects on agriculture and other industries will continue to be felt for years to come (OSBM, 2024).
  
**Why should we understand rainfall trends and why is disaster preparedness important?** 
The rainfall from Hurricane Helene was both unprecedented and catastrophic. In some areas of North Carolina, over 26 inches of rain fell within a 24-hour period, resulting in widespread flooding that devastated entire towns across 27 counties (OSBM, 2024). Analyzing precipitation patterns and discharge before, during, and after the hurricane is crucial for understanding its long-term effects on regional climates and water systems. This information is vital for informed decision-making in disaster preparedness and water resource management.
Hurricane Helene serves as a reminder for strengthening communities against flooding and extreme weather events. Climate change knows no boundaries, and we should not be surprised when such extreme events continue to occur. By focusing on Helene and its unprecedented nature, this study provides important benchmarks, analyses, and comparisons that will be invaluable for future disaster response planning.
Through examining these rainfall trends, this study aims to emphasize the historical significance of Hurricane Helene as a rare event with far-reaching consequences. Understanding these patterns offers a foundation for contextualizing the storm’s impact within the broader context of climate change, while also identifying anomalies in long-term climatic trends.


**Research Question**
*Question 1: What were the temporal and spatial rainfall patterns in Asheville, Sugar Grove, and Pigeon River following Hurricane Helene?*

*Question 2: Which counties in North Carolina were most vulnerable to extreme rainfall during and after Hurricane Helene, and why?*

*Question 3: How did urban vs. rural areas experience rainfall and flooding impacts differently?* 


**Rationale for Dataset of Choice and Data Description**
Data from the United States Geological Survey (USGS) was selected to analyze the impact and severity of Hurricane Helene, enhancing our understanding of extreme weather events. USGS stream gages collect daily discharge volumes in cubic feet per second and were used because of their accuracy and daily collection. This data analysis focused on daily precipitation trends, flood data, and hurricane patterns. To capture the diverse geographical and hydrological impacts of Helene across North Carolina, the study selected Sugar Grove, Asheville, and Pigeon River as focal sites. These locations provide a broad representation of the storm’s effects throughout Western North Carolina as they were chosen based on their proximity to Helene’s storm path. 

The study examines three critical parameters: precipitation, stream discharge, and gage height. For each parameter, daily means and cumulative sums were calculated to identify patterns and assess the storm’s severity across different regions.

FALTA (explanation on why those three: 
Sugar Grove data was used because of it's setting with agricultural production and how hard it was hit from Hurricane Helene. 
Asheville data was used due to it's size and location as a metropolitan hub in rural Western North Carolina.
Pigeon River data was used due to its proximity to the Tennessee border and its implications for the necessity of cross-border storm management strategy. 

Together, these datasets and locations provide a multi-dimensional view of Hurricane Helene’s impact, bridging historical context, storm-specific data, and hydrological consequences. This integrated approach ensures that the study not only quantifies the immediate effects of the hurricane but also contributes to long-term disaster preparedness and climate resilience.

**Data Description**

USGS stream gage data shows the latitude and longitude of the stream gage, discharge in cubic feet per second, and the date-time that the data was recorded. 
(This was already mostly in the paragraph above- we could keep this here or not)

**Describe wrangling process from raw to processed data**

MISSING 

###Table with dataset structure


```{r Data Table, echo=FALSE}
dataset_structure <- data.frame(
  Variable_Name = c(
    "Location", "Latitude", "Longitude", "DateTime", 
    "Total Daily Precipitation", "Mean Discharge", "Mean Gage Height"
  ),
  Description = c(
    "Name of the site (e.g., Asheville, Sugar Grove, Pigeon River)",
    "Latitude coordinate of the site",
    "Longitude coordinate of the site",
    "Timestamp of data observation",
    "Daily total precipitation measured at the site",
    "Mean stream discharge over the day",
    "Mean water level height above the gauge reference point"
  ),
  Data_Type = c(
    "Categorical", "Numeric", "Numeric", 
    "Date (YYYY-MM-DD)", "Numeric", "Numeric (cubic meters/sec)", 
    "Numeric (feet)"
  )
)

knitr::kable(
  dataset_structure, 
  col.names = c("Variable Name", "Description", "Data Type"),
  caption = "Dataset Structure for Analyzing Hurricane Helene's Impact",
  align = "l"
)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(warn = -1) 

library(rvest)
library(purrr)
library(lubridate)
library(dplyr)
library(zoo)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(leaflet)
library(zoo)

#install.packages("dataRetrieval")
library(dataRetrieval)

#install.packages("forecast")
library(forecast)

```


Retrieving data for Sugar Grove Gage from USGS using USGS's "dataRetrieval" package.
I find its siteNumber is 03479000 from "https://waterdata.usgs.gov/monitoring-location/03479000/#parameterCode=
00060&period=P7D&showMedian=false".


<span style="font-size:15px; font-weight:bold;">Retrieving Data</span>


**Sugar Grove**
```{r echo=TRUE, warning=FALSE}

siteNumberSugarGrove<-('03479000')

#I find the MetaData for the site
SugarGroveMeta <- whatNWISsites(siteNumbers = siteNumberSugarGrove)
print(SugarGroveMeta) 

#I then use sitedata function to look at what data is available
SugarGroveSiteData<-suppressWarnings(whatNWISdata(siteNumber=siteNumberSugarGrove))
```

Looking at what data is available, I see that the parameter codes (parmcodes)
for stream discharge (00060) , precipitation (00045), and gage height (0065).
I also see that our relevant stat codes are daily mean (0003) and sum (00006)
#

```{r message=FALSE, warning=FALSE}
pcodes<-c('00060','00045','00065') #parameters discharge, precip, and gage height
scode<-c('00003', '00006') #daily mean and daily accumulated


SugarGroveData<-readNWISdata( #dataRetrival function from USGS API
  siteNumbers = siteNumberSugarGrove,
  parameterCd= pcodes,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)
  
#View(SugarGroveData)  
```

This is our data! However, It needs to be cleaned. 
First I want to remove the 'cd' columns that indicate if the data is official or 
provisional as this won't have a significant effect on our findings. 
We can mention that in our discussion of the meta data but its not 
worth a column in our df. 

Then I want to clearly label each column with its parameter and stat code:

```{r}
CleanedDataSugarGrove<-SugarGroveData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indicate provisional or official data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
#View(CleanedDataSugarGrove)


```

I want to repeat this data retrieval and cleaning for two additional sites,
French Broad River in Asheville in Southwestern NC (03451500) 
and East Fork Pigeon River in far Southwestern NC near Canton (03456500) for
examinations of 3 different rivers in different parts of the state.

First I check the metadata and sitedata to see if similar data and parameters
are available to compare to Sugar Grove

**Asheville**

```{r, warning=FALSE}

#First the Asheville site

siteNumberAsheville<-('03451500')

#I find the MetaData for the site
AshevilleMeta <- whatNWISsites(siteNumbers = siteNumberAsheville)
print(AshevilleMeta) 

#I then use sitedata function to look at what data is available
AshevilleSiteData<-whatNWISdata(siteNumber=siteNumberAsheville)

```

Then I decide I want to create an easy data retrieval function for my chosen
parameters based on my previous retreival.

```{r}

WaterData.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
df_name<-paste0('CleanData_', siteNo)
assign(df_name, CleanData, envir= .GlobalEnv)

return(CleanData)
}
```

I run this function for Asheville's siteNo
```{r}
siteNoAsheville<-'03451500'

siteNo<-siteNoAsheville
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataAsheville<-WaterData.Function(siteNo, pcode, scode)
#View(CleanedDataAsheville)


```

I then do the same for the Pigeon River Site and its site number
**Pigeon River**

```{r}

siteNoPigeonRiver<-'03456500'

siteNo<-siteNoPigeonRiver
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataPigeonRiver<-WaterData.Function(siteNo, pcode, scode)

#View(CleanedDataPigeonRiver)
```

<span style="font-size:15px; font-weight:bold;">Research Question</span>

<span style="font-size:20px; font-weight:bold;">How unusual was Helene?</span>


First let's create separate layers from our data for 2024 and
then Helene as an event within 2024 a 7 day span from 9/22/24 to 9/29/24.
**Investigating Helene**

```{r Helene Precipitation event over 2024}

#first we create our df for 2024
Data2024_sg <- CleanedDataSugarGrove %>%
  filter(year(dateTime) == 2024)

Data2024_ash <- CleanedDataAsheville %>%
  filter(year(dateTime) == 2024)

Data2024_pg <- CleanedDataPigeonRiver %>%
  filter(year(dateTime) == 2024)




#then we isolate the Helene as a 7 day event:

HeleneData_sg <- CleanedDataSugarGrove %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))

HeleneData_ash <- CleanedDataAsheville %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))

HeleneData_pg <- CleanedDataPigeonRiver %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))


#View(HeleneData_pg)


```


Then, let's see how the week of Helene compares to the rest of 2024 within our
parameters

**Precipitation**
```{r, warning=FALSE}

# First: Sugar Grove Precipitation for 2024
ggplot(Data2024_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black", size = 2) +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black", size = 2) +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()

# Second: Asheville Precipitation for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Asheville",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()

# Third: Pigeon River Precipitation for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()




```



**Then discharge**
```{r Helene Discharge event over 2024, warning=FALSE}

# First: Sugar Grove Discharge for 2024
ggplot(Data2024_sg, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()


# Second: Asheville Discharge for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Asheville",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()




# Third: Pigeon River Discharge for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()


```



**Then Gage Height**
```{r}

# First: Sugar Grove Gage Height for 2024

ggplot(Data2024_sg, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()

# Second: Asheville Discharge for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Asheville",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()


# Third: Pigeon River Gage Height for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()




```


Okay, so clearly Helene was the maximum event for these parameters for 2024.


<span style="font-size:15px;font-style:italic;">How do these maximums compare to all historical data we have access to?</span>


Let's zoom out on the same dates but compared with all historical data.

**Precipitation**

```{r, warning=FALSE}

#First: Sugar Grive Historical Precipitation 

SugarGroveDailyPrecipitationLinePlot<-ggplot(CleanedDataSugarGrove%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x= dateTime,
   y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "lightblue", size = 1) +
  geom_point(data = HeleneData_sg, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +  
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Sugar Grove(1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 
 
print(SugarGroveDailyPrecipitationLinePlot) 


#Second: Asheville Historical Precipitation 

AshevilleDailyPrecipitationLinePlot<-ggplot(CleanedDataAsheville%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x = dateTime, 
  y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +    
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 

print(AshevilleDailyPrecipitationLinePlot)


#Third: Pigeon River Historical Precipitation 
PigeonRiverDailyTotalPrecipitationPlot<-ggplot(CleanedDataPigeonRiver%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x= dateTime,
   y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "grey", size = 1) +
  geom_point(data = HeleneData_pg, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +    
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Pigeon River (1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 

print(PigeonRiverDailyTotalPrecipitationPlot)



```


**Discharge**
```{r, warning=FALSE}

#First:Sugar Grove Historical Discharge Data

SugarGroveDailyDischargeLinePlot<-ggplot(CleanedDataSugarGrove%>%
    drop_na(`Mean Discharge`), aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="lightblue")+
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Sugar Grove(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()
 
print(SugarGroveDailyDischargeLinePlot) 



#Second: Asheville Historical Discharge Data 

AshevilleDailyDischargeLinePlot<-ggplot(CleanedDataAsheville %>%
    drop_na(`Mean Discharge`), aes(
  x = dateTime, 
  y = `Mean Discharge`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Mean Discharge (cubic meters per second)"
  ) +
  theme_minimal() 

print(AshevilleDailyDischargeLinePlot)


#Third: Pigeon River Historical Discharge Data 

PigeonRiverDailyDischargeLinePlot<-ggplot(CleanedDataPigeonRiver%>%
    drop_na(`Mean Discharge`), aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="grey")+
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Pigeon River(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()

print(PigeonRiverDailyDischargeLinePlot)


```





**Gage Height**
```{r, warning=FALSE}

#First: Sugar Grove Historical Gage Height 

SugarGroveDailyGageHeightPlot<-ggplot(CleanedDataSugarGrove %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "lightblue", size = 1) +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +    
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Sugar Grove Gage Height",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(SugarGroveDailyGageHeightPlot)


#Second: Asheville Historical Gage Height 

AshevilleDailyGageHeightPlot<-ggplot(CleanedDataAsheville %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +      
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(AshevilleDailyGageHeightPlot)


#Third: Pigeon River Historical Gage Height

PigeonRiverDailyGageHeightPlot<-ggplot(CleanedDataPigeonRiver %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "grey", size = 1) +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +      
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Pigeon River Gage Height",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(PigeonRiverDailyGageHeightPlot)


```

These graphs reaveal that Helene was responsible for some
outlier behavior, but it doesn't account for seasonality and other
trends.

Let's decompose to account for these trends
**Decompositions**

```{r SUGAR GROVE}

#Sugar Grove#
# Convert dateTime to Date class and clean columns
CleanedDataSugarGrove <- CleanedDataSugarGrove %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataSugarGrove.clean$dateTime))
s_month <- month(first(CleanedDataSugarGrove.clean$dateTime))
f_year <- year(last(CleanedDataSugarGrove.clean$dateTime))
f_month <- month(last(CleanedDataSugarGrove.clean$dateTime))

# Re-create the daily time series object
CleanedDataSugarGrove.daily.7DayAvg.ts <- ts(
  CleanedDataSugarGrove.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataSugarGrove.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataSugarGrove.daily.7DayAvg.decomp <- stl(
  CleanedDataSugarGrove.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataSugarGrove.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Sugar Grove"
)


```


```{r ASHEVILLE}

#Asheville

# Convert dateTime to Date class and clean columns
CleanedDataAsheville <- CleanedDataAsheville %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataAsheville.clean <- CleanedDataAsheville %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataAsheville.clean <- CleanedDataAsheville.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataAsheville.clean <- CleanedDataAsheville.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataAsheville.clean$dateTime))
s_month <- month(first(CleanedDataAsheville.clean$dateTime))
f_year <- year(last(CleanedDataAsheville.clean$dateTime))
f_month <- month(last(CleanedDataAsheville.clean$dateTime))

# Re-create the daily time series object
CleanedDataAsheville.daily.7DayAvg.ts <- ts(
  CleanedDataAsheville.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataAsheville.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataAsheville.daily.7DayAvg.decomp <- stl(
  CleanedDataAsheville.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataAsheville.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Asheville"
)


```


```{r PIGEON RIVER}

#Pigeon River

# Convert dateTime to Date class and clean columns
CleanedDataPigeonRiver <- CleanedDataPigeonRiver %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataPigeonRiver.clean$dateTime))
s_month <- month(first(CleanedDataPigeonRiver.clean$dateTime))
f_year <- year(last(CleanedDataPigeonRiver.clean$dateTime))
f_month <- month(last(CleanedDataPigeonRiver.clean$dateTime))

# Re-create the daily time series object
CleanedDataPigeonRiver.daily.7DayAvg.ts <- ts(
  CleanedDataPigeonRiver.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataPigeonRiver.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataPigeonRiver.daily.7DayAvg.decomp <- stl(
  CleanedDataPigeonRiver.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataPigeonRiver.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Pigeon River"
)


```


The graphs provide valuable insights, with some indicating 2024 as a statistical 
outlier. However, they do not quantify precisely how much Hurricane Helene 
deviates from historical norms. To evaluate this with statistical certainty,
we aim to determine just how unusual Helene was as an event.

To achieve this, we will apply the: Hazen Method, to calculate the recurrence 
intervals for Hurricane Helene within the analyzed parameters. 
This approach will help contextualize the storm’s magnitude and frequency 
compared to historical data.


First we will calculate the recurrence intervals for 7-day precipitation events. 
Based on our initial analysis, assuming Hurricane Helene represents the 7-day 
averaged maximum for 2024 is a reasonable starting point.



<span style="font-size:18px; font-weight:bold;">Helene Recurrence Intervals</span>

**Sugar Grove Hazen Precipitation**

```{r Sugar Grove Precip Hazen, warning=FALSE}
SugarGrovePrecip <- CleanedDataSugarGrove %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

SugarGrovePrecip <- SugarGrovePrecip %>%
  mutate(Year = year(dateTime))

SugarGrovePrecip <- SugarGrovePrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))

SugarGrovePrecip <- SugarGrovePrecip %>% 
  filter(MaxEvent > 0)

SugarGrovePrecip <- SugarGrovePrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

SugarGrovePrecipHazen2024 <- SugarGrovePrecip %>% filter(Year == 2024)
print(SugarGrovePrecipHazen2024)


```

The Hazen number of 18 indicates that the maximum 7-day average precipitation 
event in 2024 was an 18-year event, meaning it had about a 20% chance of 
occurring in any given year (Hazen interval: 18 years, probability: 20%).

**Asheville Hazen Precipitation**
```{r Asheville Precipitation Hazen, warning=FALSE}
AshevillePrecip <- CleanedDataAsheville %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

AshevillePrecip <- AshevillePrecip %>%
  mutate(Year = year(dateTime))

AshevillePrecip <- AshevillePrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm = TRUE))

AshevillePrecip <- AshevillePrecip %>% 
  filter(MaxEvent > 0)

AshevillePrecip <- AshevillePrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100 * (2 * rank - 1)) / (2 * n()))) %>% 
  mutate(hazen = 100 / Fa_Value)

AshevillePrecipHazen2024 <- AshevillePrecip %>% filter(Year == 2024)
print(AshevillePrecipHazen2024)


```

The Hazen number reveals that the 7-day average discharge observed in Asheville
in 2024 qualifies as a 54-year event. In other words, such an event is expected 
to occur approximately once every 54 years, equating to a 2% probability of 
happening in any given year.  (Hazen interval: 54 years, probability: 2%)

This provides a statistical metric to quantify the observed outliers. 

Next, we will calculate the recurrence intervals for 2024 at our last 
selected site.

**Pigeon River Hazen Precipitation**

```{r Pigeon River Precip Hazen, warning=FALSE}
PigeonRiverPrecip <- CleanedDataPigeonRiver %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

PigeonRiverPrecip <- PigeonRiverPrecip %>%
  mutate(Year = year(dateTime))

PigeonRiverPrecip <- PigeonRiverPrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm = TRUE))

PigeonRiverPrecip <- PigeonRiverPrecip  %>% 
  filter(MaxEvent > 0)

PigeonRiverPrecip <- PigeonRiverPrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100 * (2 * rank - 1)) / (2 * n()))) %>% 
  mutate(hazen = 100 / Fa_Value)

PigeonRiverPrecipHazen2024 <- PigeonRiverPrecip %>% filter(Year == 2024)

print(PigeonRiverPrecipHazen2024)
```

At Pigeon River, Hurricane Helene was a 52-year event in terms of total
precipitation. This indicates that Helene's precipitation was less unusual at 
Sugar Grove than it was in Asheville or Pigeon River during the
maximum 7-day event. (Hazen interval: 52 years, probability: ~2%)


Let's now automate this function for all three parameters and run our sites.


**Automated Precipitation**
```{r}

Hazen2024Precip.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
CleanDataFlood<-CleanData%>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

CleanDataFlood<- CleanDataFlood%>%
  mutate(Year = year(dateTime))

CleanDataFlood <- CleanDataFlood%>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))

CleanDataFlood <- CleanDataFlood %>% 
  filter(MaxEvent > 0)

CleanDataFlood <- CleanDataFlood %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

CleanDataFlood2024 <- CleanDataFlood %>% filter(Year == 2024)


return(CleanDataFlood2024)
}

```


**Automated Discharge**
```{r}

Hazen2024Discharge.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
CleanDataFlood<-CleanData%>% 
  mutate(moving_avg = rollapply(`Mean Discharge`, width = 7, FUN = mean, align = "center", fill = NA)) 

CleanDataFlood<- CleanDataFlood%>%
  mutate(Year = year(dateTime))

CleanDataFlood <- CleanDataFlood%>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))

CleanDataFlood <- CleanDataFlood %>% 
  filter(MaxEvent > 0)

CleanDataFlood <- CleanDataFlood %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

CleanDataFlood2024 <- CleanDataFlood %>% filter(Year == 2024)


return(CleanDataFlood2024)
}

```

*And now we run our sites through our discharge function*

```{r}

SugarGroveDischargeHazen2024 <- Hazen2024Discharge.Function(siteNumberSugarGrove, pcode, scode)

AshevilleDischargeHazen2024 <- Hazen2024Discharge.Function(siteNumberAsheville, pcode, scode)

PigeonRiverDischargeHazen2024 <- Hazen2024Discharge.Function(siteNoPigeonRiver, pcode, scode)

SugarGroveDischargeHazen2024
AshevilleDischargeHazen2024
PigeonRiverDischargeHazen2024



```

*Now our function for Gage Height*

```{r}
Hazen2024GageHeight.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
CleanDataGageHeight<-CleanData%>% 
  mutate(moving_avg = rollapply(`Mean Gage Height`, width = 7, FUN = mean, align = "center", fill = NA)) 

CleanDataGageHeight<- CleanDataGageHeight%>%
  mutate(Year = year(dateTime))

CleanDataGageHeight <- suppressWarnings(CleanDataGageHeight%>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))
)

CleanDataGageHeight <- CleanDataGageHeight %>% 
  filter(MaxEvent > 0)

CleanDataGageHeight <- CleanDataGageHeight %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

CleanDataGageHeight2024 <- CleanDataGageHeight %>% filter(Year == 2024)


return(CleanDataGageHeight2024)
}

#testing function

SugarGroveGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNumberSugarGrove, pcode, scode)

AshevilleGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNumberAsheville, pcode, scode)

PigeonRiverGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNoPigeonRiver, pcode, scode)

SugarGroveGageHeightHazen2024
AshevilleGageHeightHazen2024
PigeonRiverGageHeightHazen2024

```



**Map of Gage Sites**

To visually illustrate how unusual Hurricane Helene was at each site,
we will create a map that represents the magnitude of the Hazen interval. 

The size of the dot on the map will correspond to the rarity of the event: 
the larger the dot, the less likely the event was to occur.
This visualization will help convey the statistical significance of 
Helene's impact across the selected sites.

```{r}



# Create Data Frame with Hazen Values
hazen_data <- data.frame(
  Location = c("Pigeon River", "Asheville", "Sugar Grove"),
  Latitude = c(35.554, 35.595, 36.267),
  Longitude = c(-82.982, -82.551, -81.785),
  Precipitation_Hazen = sapply(list(PigeonRiverPrecipHazen2024$hazen, 
                                    AshevillePrecipHazen2024$hazen, 
                                    SugarGrovePrecipHazen2024$hazen), function(x) as.numeric(unlist(x))),
  Discharge_Hazen = sapply(list(PigeonRiverDischargeHazen2024$hazen, 
                                 AshevilleDischargeHazen2024$hazen, 
                                 SugarGroveDischargeHazen2024$hazen), function(x) as.numeric(unlist(x))),
  GageHeight_Hazen = sapply(list(PigeonRiverGageHeightHazen2024$hazen, 
                                  AshevilleGageHeightHazen2024$hazen, 
                                  SugarGroveGageHeightHazen2024$hazen), function(x) as.numeric(unlist(x)))
)


# Define distinct colors for each layer
precip_color <- "blue"
discharge_color <- "green"
gage_height_color <- "orange"

# Define a scaling factor for marker sizes
base_size <- 3  # Base size for markers
scaling_factor <- 10  # Scaling factor for marker sizes

# Create a custom function to add size legend
addLegendSize <- function(map, title, sizes, values, colors, position = "bottomright") {
  size_labels <- paste0(values)
  legend_html <- paste0(
    "<div style='padding: 5px;'><strong>", title, "</strong><br>",
    paste0(
      mapply(function(size, label) {
        paste0(
          "<div style='display: flex; align-items: center;'>",
          "<div style='height:", size * 2, "px; width:", size * 2, "px; ",
          "border-radius: 50%; background-color:", colors, "; ",
          "margin-right: 8px;'></div>",
          label, "</div>"
        )
      }, sizes, size_labels, SIMPLIFY = FALSE),
      collapse = ""
    ),
    "</div>"
  )
  addControl(map, html = legend_html, position = position)
}

# Predefined Hazen value examples for the size legend
size_values <- c(10, 50, 100, 200)  # Example Hazen values
size_sizes <- base_size + size_values / scaling_factor  # Corresponding marker sizes

# Create the Leaflet Map
leaflet(hazen_data) %>%
  addTiles() %>%
  
  # Precipitation Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = precip_color,
    fillColor = precip_color,
    fillOpacity = 1,   # Ensure full opacity
    radius = ~base_size + Precipitation_Hazen / scaling_factor,  # Scale marker size
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Precipitation Hazen: ", round(Precipitation_Hazen, 2)
    ),
    group = "Precipitation"
  ) %>%
  
  # Discharge Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = discharge_color,
    fillColor = discharge_color,
    fillOpacity = 1,   # Ensure full opacity
    radius = ~base_size + Discharge_Hazen / scaling_factor,  # Scale marker size
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Discharge Hazen: ", round(Discharge_Hazen, 2)
    ),
    group = "Discharge"
  ) %>%
  
  # Gage Height Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = gage_height_color,
    fillColor = gage_height_color,
    fillOpacity = 1,   # Ensure full opacity
    radius = ~base_size + GageHeight_Hazen / scaling_factor,  # Scale marker size
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Gage Height Hazen: ", round(GageHeight_Hazen, 2)
    ),
    group = "Gage Height"
  ) %>%
  
  # Add Layer Controls
  addLayersControl(
    overlayGroups = c("Precipitation", "Discharge", "Gage Height"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  
  # Add a Color Legend for Layers
  addLegend(
    "bottomleft",
    colors = c(precip_color, discharge_color, gage_height_color),
    labels = c("Precipitation", "Discharge", "Gage Height"),
    title = "Hazen Value Layers",
    opacity = 1
  ) %>%
  
  # Add a Size Legend for Hazen Values
  addLegendSize(
    title = "Hazen Value (Size)",
    sizes = size_sizes,
    values = size_values,
    colors = "grey",
    position = "bottomright"
  )



```

What we found was very interesting! 
Now that we have an easy function, we will download more sites in Western North
Carolina that have precipitation, discharge, and gage height available. This will
be a helpful visual in understanding the specific impacts in Sugar Grove, Asheville, 
and Pigeon River compared to the rest of the state. 


```{r message=FALSE}

#CHEOAH RIVER NR BEARPEN GAP NR TAPOCO, NC 
CheoahRiverSiteNo<-as.character('0351706800')
CheoahRiverGageHeightHazen2024<-Hazen2024GageHeight.Function(CheoahRiverSiteNo , pcode, scode)
CheoahRiverDischargeHazen2024<-Hazen2024Discharge.Function(CheoahRiverSiteNo , pcode, scode)
CheoahRiverPrecipHazen2024<-Hazen2024Precip.Function(CheoahRiverSiteNo , pcode, scode)

#SWANNANOA RIVER AT BILTMORE, NC 
SwannanoaSiteNo<-as.character('03451000')
SwannanoaGageHeightHazen2024<-Hazen2024GageHeight.Function(SwannanoaSiteNo, pcode, scode)
SwannanoaDischargeHazen2024<-Hazen2024Discharge.Function(SwannanoaSiteNo, pcode, scode)
SwannanoaPrecipHazen2024<-Hazen2024Precip.Function(SwannanoaSiteNo, pcode, scode)

#SECOND BROAD RIVER NR LOGAN, NC
SecondBroadRiverSiteNo<-as.character('02150495')
SecondBroadRiverGageHeightHazen2024<-Hazen2024GageHeight.Function(SecondBroadRiverSiteNo , pcode, scode)
SecondBroadRiverDischargeHazen2024<-Hazen2024Discharge.Function(SecondBroadRiverSiteNo , pcode, scode)
SecondBroadRiverPrecipHazen2024<-Hazen2024Precip.Function(SecondBroadRiverSiteNo, pcode, scode)

#CHEOAH RIVER NR BEARPEN GAP NR TAPOCO, NC 
CheoahRiverSiteNo<-as.character('0351706800')
CheoahGageHeightHazen2024<-Hazen2024GageHeight.Function(CheoahRiverSiteNo , pcode, scode)
CheoahDischargeHazen2024<-Hazen2024Discharge.Function(CheoahRiverSiteNo , pcode, scode)
CheoahPrecipHazen2024<-Hazen2024Precip.Function(CheoahRiverSiteNo , pcode, scode)

#ELK CREEK AT ELKVILLE, NC
ElkCreekSiteNo<-as.character('02111180')
ElkCreekGageHeightHazen2024<-Hazen2024GageHeight.Function(ElkCreekSiteNo , pcode, scode)
ElkCreekDischargeHazen2024<-Hazen2024Discharge.Function(ElkCreekSiteNo , pcode, scode)
ElkCreekPrecipHazen2024<-Hazen2024Precip.Function(ElkCreekSiteNo, pcode, scode)

#TAR RIVER AT NC 97 AT ROCKY MOUNT, NC 
TarRiverSiteNo<-as.character('02082585')
TarRiverGageHeightHazen2024<-Hazen2024GageHeight.Function(TarRiverSiteNo , pcode, scode)
TarRiverDischargeHazen2024<-Hazen2024Discharge.Function(TarRiverSiteNo , pcode, scode)
TarRiverPrecipHazen2024<-Hazen2024Precip.Function(TarRiverSiteNo , pcode, scode)


```


After selecting the additional sites, they were Incorporated into an array to 
enhance the map. 
The map now displays the magnitude of each Hazen interval, 
representing how unusual the 7-day Hurricane Helene event was at each site. 
The size of the dot corresponds to the rarity of the event:
larger dots indicate more unusual 7-day averages for the given parameter. 
Although the number of sites with all three parameters is limited, 
further analysis could enable the creation of increasingly granular data for 
each site, improving the resolution of our insights.




```{r}

# Create Data Frame with Hazen Values
Eight_site_hazen_data <- data.frame(
  Location = c(
    "Pigeon River", "Asheville", "Sugar Grove", 
    "Cheoah River", "Swannanoa River", "Second Broad River", 
    "Elk Creek", "Tar River"
  ),
  Latitude = c(
    35.554, 35.595, 36.267, 
    35.447, 35.572, 35.362, 
    36.153, 35.942
  ),
  Longitude = c(
    -82.982, -82.551, -81.785, 
    -83.852, -82.544, -81.958, 
    -81.421, -77.814
  ),
  Precipitation_Hazen = sapply(list(
    PigeonRiverPrecipHazen2024$hazen, AshevillePrecipHazen2024$hazen, SugarGrovePrecipHazen2024$hazen,
    CheoahRiverPrecipHazen2024$hazen, SwannanoaPrecipHazen2024$hazen, SecondBroadRiverPrecipHazen2024$hazen,
    ElkCreekPrecipHazen2024$hazen, TarRiverPrecipHazen2024$hazen
  ), function(x) as.numeric(unlist(x))),
  Discharge_Hazen = sapply(list(
    PigeonRiverDischargeHazen2024$hazen, AshevilleDischargeHazen2024$hazen, SugarGroveDischargeHazen2024$hazen,
    CheoahRiverDischargeHazen2024$hazen, SwannanoaDischargeHazen2024$hazen, SecondBroadRiverDischargeHazen2024$hazen,
    ElkCreekDischargeHazen2024$hazen, TarRiverDischargeHazen2024$hazen
  ), function(x) as.numeric(unlist(x))),
  GageHeight_Hazen = sapply(list(
    PigeonRiverGageHeightHazen2024$hazen, AshevilleGageHeightHazen2024$hazen, SugarGroveGageHeightHazen2024$hazen,
    CheoahRiverGageHeightHazen2024$hazen, SwannanoaGageHeightHazen2024$hazen, SecondBroadRiverGageHeightHazen2024$hazen,
    ElkCreekGageHeightHazen2024$hazen, TarRiverGageHeightHazen2024$hazen
  ), function(x) as.numeric(unlist(x)))
)

# Define distinct colors for each layer
precip_color <- "blue"
discharge_color <- "green"
gage_height_color <- "orange"

# Define a scaling factor for marker sizes
base_size <- 3  # Base size for markers
scaling_factor <- 10  # Scaling factor for marker sizes

# Create a custom function to add size legend
addLegendSize <- function(map, title, sizes, values, colors, position = "bottomright") {
  size_labels <- paste0(values)
  legend_html <- paste0(
    "<div style='padding: 5px;'><strong>", title, "</strong><br>",
    paste0(
      mapply(function(size, label) {
        paste0(
          "<div style='display: flex; align-items: center;'>",
          "<div style='height:", size * 2, "px; width:", size * 2, "px; ",
          "border-radius: 50%; background-color:", colors, "; ",
          "margin-right: 8px;'></div>",
          label, "</div>"
        )
      }, sizes, size_labels, SIMPLIFY = FALSE),
      collapse = ""
    ),
    "</div>"
  )
  addControl(map, html = legend_html, position = position)
}

# Predefined Hazen value examples for the size legend
size_values <- c(10, 50, 100, 200)  # Example Hazen values
size_sizes <- base_size + size_values / scaling_factor  # Corresponding marker sizes

# Create the Leaflet Map
leaflet(Eight_site_hazen_data) %>%
  addTiles() %>%
  
  # Precipitation Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = precip_color,
    fillColor = precip_color,
    fillOpacity = 1,   # Ensure full opacity
    radius = ~base_size + Precipitation_Hazen / scaling_factor,  # Scale marker size
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Precipitation Hazen: ", round(Precipitation_Hazen, 2)
    ),
    group = "Precipitation"
  ) %>%
  
  # Discharge Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = discharge_color,
    fillColor = discharge_color,
    fillOpacity = 1,   # Ensure full opacity
    radius = ~base_size + Discharge_Hazen / scaling_factor,  # Scale marker size
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Discharge Hazen: ", round(Discharge_Hazen, 2)
    ),
    group = "Discharge"
  ) %>%
  
  # Gage Height Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = gage_height_color,
    fillColor = gage_height_color,
    fillOpacity = 1,   # Ensure full opacity
    radius = ~base_size + GageHeight_Hazen / scaling_factor,  # Scale marker size
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Gage Height Hazen: ", round(GageHeight_Hazen, 2)
    ),
    group = "Gage Height"
  ) %>%
  
  # Add Layer Controls
  addLayersControl(
    overlayGroups = c("Precipitation", "Discharge", "Gage Height"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  
  # Add a Color Legend for Layers
  addLegend(
    "bottomleft",
    colors = c(precip_color, discharge_color, gage_height_color),
    labels = c("Precipitation", "Discharge", "Gage Height"),
    title = "Hazen Value Layers",
    opacity = 1
  ) %>%
  
  # Add a Size Legend for Hazen Values
  addLegendSize(
    title = "Hazen Value (Size)",
    sizes = size_sizes,
    values = size_values,
    colors = "grey",
    position = "bottomright"
  )






```
**Why this analysis matters**

This map can be understood as a representation of relative intensity compared
to historical norms. While the size of the dots does not directly indicate the
absolute intensity of precipitation or flooding, it reflects the magnitude of 
these parameters relative to each area's historical averages.
Flood management infrastructure is typically designed to handle 
50-year or 75-year storms, making the relative magnitude to historical norms 
critical for assessing how well the infrastructure performed under 
Hurricane Helene's conditions—or how severely it was overwhelmed.
Although the number of sites with all three parameters is limited,
the analysis reveals that Helene's relative intensity to historical 
norms was greatest along a path through central Western North Carolina, 
particularly over Asheville. From this, we can reasonably infer that Asheville’s 
infrastructure experienced greater inundation compared to areas in northeastern 
North Carolina or other sites included in this study.
