---
title: "Final Project Draft"
author: "Sam Tolbert, Chrissie Pantoja, Ana Andino, Gretchen Barbera"
date: "2024-12-09"
output:
  html_document:
    toc: true
    toc_float: true
---
##Context and Motivation
Hurricane Helene struck North Carolina in late September after it quickly intensified to a Category 4 hurricane with sustained winds of 140 mph. Hurricane Helene brought historical rainfall, catastrophic flooding, landslides, strong winds and tornadoes that struck the economy and livelihoods in such a way that up to this date, they are still recovering. Most striking is that Helene happened just a few days after Hurricane Milton (category 5) had also hit the west coast of the US.  These events underscore the growing severity and frequency of hurricanes, highlighting the urgent need to understand the mechanisms and impacts of such extreme weather events.

 Hurricane Helene has left a profound and lasting impact on North Carolina. The storm caused unprecedented destruction, leaving thousands of homes destroyed or damaged. Millions of residents were left without access to critical infrastructure, including water, electricity, telecommunications, and healthcare (OSBM, 2024). Infrastructure such as roadways and bridges were severely damaged. In economic terms, the impact of the storm will be long standing. Preliminary estimates of damage and needs across the state approach $53.6 billion, including $41.1 billion in direct damage, $7.6 billion in indirect impacts, and $4.8 billion in necessary investments for strengthening and mitigation (OSBM, 2024). These figures highlight the storm's disproportionate impact compared to other recent hurricanes, such as Hurricane Florence in 2018, whose total damages amounted to $16.7 billion. Beyond economic losses, the storm reshaped the physical landscape eroding valuable topsoil, altering river paths, and submerging homes that had stood for generations under newly formed waterways plus affecting agriculture production in the long run as well (OSBM, 2024).
  
###Why should we understand rainfall trends and why is disaster preparedness important? 
The rainfall brought by Hurricane Helene was both unprecedented and catastrophic. Some areas in North Carolina received over 26 inches of rain within a 24-hour period, leading to flooding that decimated entire towns across 27 counties (OSBM, 2024). Analyzing precipitation patterns and discharge before, during, and after the hurricane is critical for understanding its lingering effects on regional climates and water systems. This is necessary to make inform decisions in disaster preparedness and water resource management.  
Hurricane Helene now serves as a reminder of the importance of strengthening communities against flooding and extreme weather. Climate change has no boundaries, and we should be surprised of these extreme effects keep happening thus by focusing on Helene and its uniqueness, the study provides important benchmarks, analysis and comparisons to think about future disaster response.
By examining these rainfall trends, this study seeks to highlight the unique nature of Hurricane Helene historically as a rare phenomenon with broad consequences. Understanding these patterns provides a foundation for contextualizing Helene's impact within the wider scope of climate change while identifying anomalies in long-term climatic trends. 

##Research Question 
Question 1: What were the temporal and spatial rainfall patterns in the countisof Asheville, Sugar Grove, and Pigeon River following Hurricane Helene? 

Question 2: Which counties in North Carolina were most vulnerable to extreme rainfall during and after Hurricane Helene, and why? 

Question 3: How did urban vs. rural areas experience rainfall and flooding impacts differently? 


###Rationale for Dataset of Choice 
Data from the United States Geological Survey (USGS) was selected to analyze the impact and severity of Hurricane Helene and to better understanding extreme weather events. To do so, we focused on daily precipitation trends, flood data and hurricane patterns. Additionally, the selection of Asheville, Sugar Grove, and Pigeon River as focal sites ensures both geographical and hydrological diversity, capturing a wide range of storm impacts in North Carolina. The study focuses in three critical parameters: Stream Discharge, Precipitation, and Gage Height. For them, we find the daily means and sums

FALTA (explanation on why those three: Asheville is a small urban place that is typically not affected by these events, sugar grove a setting with agro production and pigeon river ??)

Together, these datasets and locations provide a multi-dimensional view of Hurricane Helene’s impact, bridging historical context, storm-specific data, and hydrological consequences. This integrated approach ensures that the study not only quantifies the immediate effects of the hurricane but also contributes to long-term disaster preparedness and climate resilience.
##Data Description 

(Data suources and content)

##Describe wrangling process from raw to processed data 

MISSING 
###Table with datsdet structure


```{r Data Table, echo=FALSE}
dataset_structure <- data.frame(
  Variable_Name = c(
    "Location", "Latitude", "Longitude", "DateTime", 
    "Total Daily Precipitation", "Mean Discharge", "Mean Gage Height"
  ),
  Description = c(
    "Name of the site (e.g., Asheville, Sugar Grove, Pigeon River)",
    "Latitude coordinate of the site",
    "Longitude coordinate of the site",
    "Timestamp of data observation",
    "Daily total precipitation measured at the site",
    "Mean stream discharge over the day",
    "Mean water level height above the gauge reference point"
  ),
  Data_Type = c(
    "Categorical", "Numeric", "Numeric", 
    "Date (YYYY-MM-DD)", "Numeric", "Numeric (cubic meters/sec)", 
    "Numeric (feet)"
  )
)

kable(
  dataset_structure, 
  col.names = c("Variable Name", "Description", "Data Type"),
  caption = "Dataset Structure for Analyzing Hurricane Helene's Impact",
  align = "l"
)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rvest)
library(purrr)
library(lubridate)
library(here)
library(dplyr)
library(zoo)
library(ggplot2)
library(tidyr)

#install.packages("dataRetrieval")
library(dataRetrieval)

#install.packages("forecast")
library(forecast)

```


Retrieving data for Sugar Grove Gage from USGS using USGS's "dataRetrieval" package.
I find its siteNumber is 03479000 from "https://waterdata.usgs.gov/monitoring-location/03479000/#parameterCode=
00060&period=P7D&showMedian=false".


<span style="font-size:15px; font-weight:bold;">Retrieving Data</span>


**Sugar Grove**
```{r echo=TRUE, warning=FALSE}

siteNumberSugarGrove<-('03479000')

#I find the MetaData for the site
SugarGroveMeta <- whatNWISsites(siteNumbers = siteNumberSugarGrove)
print(SugarGroveMeta) 

#I then use sitedata function to look at what data is available
SugarGroveSiteData<-suppressWarnings(whatNWISdata(siteNumber=siteNumberSugarGrove))
```

Looking at what data is available, I see that the parameter codes (parmcodes)
for stream discharge (00060) , precipitation (00045), and gage height (0065).
I also see that our relevant stat codes are daily mean (0003) and sum (00006)
#

```{r message=FALSE, warning=FALSE}
pcodes<-c('00060','00045','00065') #parameters discharge, precip, and gage height
scode<-c('00003', '00006') #daily mean and daily accumulated


SugarGroveData<-readNWISdata( #dataRetrival function from USGS API
  siteNumbers = siteNumberSugarGrove,
  parameterCd= pcodes,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)
  
#View(SugarGroveData)  
```

This is my data! However, I need to clean this up. 
First I want to remove the 'cd' columns that indicate if the data is official or 
provisional as this won't have a significant effect on our findings. 
We can mention that in our discussion of the meta data but its not 
worth a column in our df. 

Then I want to clearly label each column with its parameter and stat code:

```{r}
CleanedDataSugarGrove<-SugarGroveData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
#View(CleanedDataSugarGrove)


```

I want to repeat this data retrieval and cleaning for two additional sites,
French Broad River in Asheville in Southwestern NC (03451500) 
and East Fork Pigeon River in far Southwestern NC near Canton (03456500) for
examinations of 3 different rivers in different parts of the state.

First I check the metadata and sitedata to see if similar data and parameters
are availale to compare to Sugar Grove

**Asheville**

```{r, warning=FALSE}

#First the Asheville site

siteNumberAsheville<-('03451500')

#I find the MetaData for the site
AshevilleMeta <- whatNWISsites(siteNumbers = siteNumberAsheville)
print(AshevilleMeta) 

#I then use sitedata function to look at what data is available
AshevilleSiteData<-whatNWISdata(siteNumber=siteNumberAsheville)

```

Then I decide I want to create an easy data retrieval function for my chosen
parameters based on my previous retreival.

```{r}

WaterData.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
df_name<-paste0('CleanData_', siteNo)
assign(df_name, CleanData, envir= .GlobalEnv)

return(CleanData)
}
```

I run this function for Asheville's siteNo
```{r}
siteNoAsheville<-'03451500'

siteNo<-siteNoAsheville
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataAsheville<-WaterData.Function(siteNo, pcode, scode)
#View(CleanedDataAsheville)


```

I then do the same for the Pigeon River Site and its site number
**Pigeon River**

```{r}

siteNoPigeonRiver<-'03456500'

siteNo<-siteNoPigeonRiver
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataPigeonRiver<-WaterData.Function(siteNo, pcode, scode)

#View(CleanedDataPigeonRiver)
```

<span style="font-size:15px; font-weight:bold;">Research Question</span>

<span style="font-size:20px; font-weight:bold;">How unusual was Helene?</span>


First let's create separate layers from our data for 2024 and
then Helene as an event within 2024 a 7 day span from 9/22/24 to 9/29/24.
**Investigating Helene**

```{r Helene Precipitation event over 2024}

#first we create our df for 2024
Data2024_sg <- CleanedDataSugarGrove %>%
  filter(year(dateTime) == 2024)

Data2024_ash <- CleanedDataAsheville %>%
  filter(year(dateTime) == 2024)

Data2024_pg <- CleanedDataPigeonRiver %>%
  filter(year(dateTime) == 2024)




#then we isolate the Helene as a 7 day event:

HeleneData_sg <- CleanedDataSugarGrove %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))

HeleneData_ash <- CleanedDataAsheville %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))

HeleneData_pg <- CleanedDataPigeonRiver %>%
  filter(dateTime >= as.Date("2024-09-22") & dateTime <= as.Date("2024-09-29"))


#View(HeleneData_pg)


```


Then, let's see how the week of Helene compares to the rest of 2024 within our
parameters

**Precipitation**
```{r, warning=FALSE}

# First: Sugar Grove Precipitation for 2024
ggplot(Data2024_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black", size = 2) +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black", size = 2) +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()

# Second: Asheville Precipitation for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Asheville",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()

# Third: Pigeon River Precipitation for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +
  labs(
    title = "Precipitation Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal()




```



**Then discharge**
```{r Helene Discharge event over 2024, warning=FALSE}

# First: Sugar Grove Discharge for 2024
ggplot(Data2024_sg, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()


# Second: Asheville Discharge for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Asheville",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()




# Third: Pigeon River Discharge for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Mean Discharge`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Discharge Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Discharge"
  ) +
  theme_minimal()


```



**Then Gage Height**
```{r}

# First: Sugar Grove Gage Height for 2024

ggplot(Data2024_sg, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "lightblue") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Sugar Grove",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()

# Second: Asheville Discharge for 2024
ggplot(Data2024_ash, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "darkred") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Asheville",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()


# Third: Pigeon River Gage Height for 2024
ggplot(Data2024_pg, aes(x = dateTime, y = `Mean Gage Height`)) +
  geom_line(color = "grey") +
  geom_vline(xintercept = as.Date("2024-09-24"), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.Date("2024-09-27"), linetype = "dashed", color = "black") +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +
  labs(
    title = "Gage Height Trends for 2024 in Pigeon River",
    x = "Date",
    y = "Gage Height"
  ) +
  theme_minimal()




```


Okay, so clearly Helen was the maximum event for these parameters for 2024.


<span style="font-size:15px;font-style:italic;">How do these maximums compare to all historical data we have access to?</span>


Let's zoom out on the same dates but compared with all historical data.

**Precipitation**

```{r, warning=FALSE}

#First: Sugar Grive Historical Precipitation 

SugarGroveDailyPrecipitationLinePlot<-ggplot(CleanedDataSugarGrove%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x= dateTime,
   y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "lightblue", size = 1) +
  geom_point(data = HeleneData_sg, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +  
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Sugar Grove(1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 
 
print(SugarGroveDailyPrecipitationLinePlot) 


#Second: Asheville Historical Precipitation 

AshevilleDailyPrecipitationLinePlot<-ggplot(CleanedDataAsheville%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x = dateTime, 
  y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +    
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 

print(AshevilleDailyPrecipitationLinePlot)


#Third: Pigeon River Historical Precipitation 
PigeonRiverDailyTotalPrecipitationPlot<-ggplot(CleanedDataPigeonRiver%>%
    drop_na(`Total Daily Precipitation (Inches)`), aes(
  x= dateTime,
   y = `Total Daily Precipitation (Inches)`)) +
  geom_line(color = "grey", size = 1) +
  geom_point(data = HeleneData_pg, aes(x = dateTime,
                                      y = `Total Daily Precipitation (Inches)`),
             color = "red",
             size = 3) +    
  labs(
    title = "Total Daily Precipitation Over Time",
    subtitle = "Pigeon River (1954-2024)",
    x = "Date",
    y = "Total Daily Precipitation (Inches)"
  ) +
  theme_minimal() 

print(PigeonRiverDailyTotalPrecipitationPlot)



```


**Discharge**
```{r, warning=FALSE}

#First:Sugar Grove Historical Discharge Data

SugarGroveDailyDischargeLinePlot<-ggplot(CleanedDataSugarGrove%>%
    drop_na(`Mean Discharge`), aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="lightblue")+
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Sugar Grove(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()
 
print(SugarGroveDailyDischargeLinePlot) 



#Second: Asheville Historical Discharge Data 

AshevilleDailyDischargeLinePlot<-ggplot(CleanedDataAsheville %>%
    drop_na(`Mean Discharge`), aes(
  x = dateTime, 
  y = `Mean Discharge`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Mean Discharge (cubic meters per second)"
  ) +
  theme_minimal() 

print(AshevilleDailyDischargeLinePlot)


#Third: Pigeon River Historical Discharge Data 

PigeonRiverDailyDischargeLinePlot<-ggplot(CleanedDataPigeonRiver%>%
    drop_na(`Mean Discharge`), aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="grey")+
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Discharge`),
             color = "red",
             size = 3) +
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Pigeon River(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()

print(PigeonRiverDailyDischargeLinePlot)


```





**Gage Height**
```{r, warning=FALSE}

#First: Sugar Grove Historical Gage Height 

SugarGroveDailyGageHeightPlot<-ggplot(CleanedDataSugarGrove %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "lightblue", size = 1) +
  geom_point(data = HeleneData_sg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +    
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Sugar Grove Gage Height",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(SugarGroveDailyGageHeightPlot)


#Second: Asheville Historical Gage Height 

AshevilleDailyGageHeightPlot<-ggplot(CleanedDataAsheville %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "darkred", size = 1) +
  geom_point(data = HeleneData_ash, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +      
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(AshevilleDailyGageHeightPlot)


#Third: Pigeon River Historical Gage Height

PigeonRiverDailyGageHeightPlot<-ggplot(CleanedDataPigeonRiver %>%
    drop_na(`Mean Gage Height`), aes(
  x = dateTime, 
  y = `Mean Gage Height`)) +
  geom_line(color = "grey", size = 1) +
  geom_point(data = HeleneData_pg, aes(x = dateTime, y = `Mean Gage Height`),
             color = "red",
             size = 3) +      
  labs(
    title = "Mean Daily Gage Height Over Time",
    subtitle = "Pigeon River Gage Height",
    x = "Date",
    y = "Mean Gage Height"
  ) +
  theme_minimal() 

print(PigeonRiverDailyGageHeightPlot)


```

These graphs reaveal that Helene was responsible for some
outlier behavior, but it doesn't account for seasonality and other
trends.

Let's decompose to account for these trends

###CHRISSIE PUT YOUR DECOMPS HERE###

```{r SUGAR GROVE}


# Convert dateTime to Date class and clean columns
CleanedDataSugarGrove <- CleanedDataSugarGrove %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataSugarGrove.clean <- CleanedDataSugarGrove.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataSugarGrove.clean$dateTime))
s_month <- month(first(CleanedDataSugarGrove.clean$dateTime))
f_year <- year(last(CleanedDataSugarGrove.clean$dateTime))
f_month <- month(last(CleanedDataSugarGrove.clean$dateTime))

# Re-create the daily time series object
CleanedDataSugarGrove.daily.7DayAvg.ts <- ts(
  CleanedDataSugarGrove.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataSugarGrove.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataSugarGrove.daily.7DayAvg.decomp <- stl(
  CleanedDataSugarGrove.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataSugarGrove.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Sugar Grove"
)


```


```{r}
# Convert dateTime to Date class and clean columns
CleanedDataAsheville <- CleanedDataAsheville %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataAsheville.clean <- CleanedDataAsheville %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataAsheville.clean <- CleanedDataAsheville.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataAsheville.clean <- CleanedDataAsheville.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataAsheville.clean$dateTime))
s_month <- month(first(CleanedDataAsheville.clean$dateTime))
f_year <- year(last(CleanedDataAsheville.clean$dateTime))
f_month <- month(last(CleanedDataAsheville.clean$dateTime))

# Re-create the daily time series object
CleanedDataAsheville.daily.7DayAvg.ts <- ts(
  CleanedDataAsheville.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataAsheville.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataAsheville.daily.7DayAvg.decomp <- stl(
  CleanedDataAsheville.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataAsheville.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Asheville"
)


```


```{r PIGEON RIVER}
library(dplyr)
library(lubridate)
library(zoo)


# Convert dateTime to Date class and clean columns
CleanedDataPigeonRiver <- CleanedDataPigeonRiver %>%
  mutate(
    dateTime = as.Date(dateTime),
    mean_discharge = as.numeric(`Mean Discharge`),
    `Mean Gage Height` = as.numeric(`Mean Gage Height`),
    `Total Daily Precipitation (Inches)` = as.numeric(`Total Daily Precipitation (Inches)`)
  )

# Fill missing Mean Discharge values using linear interpolation
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver %>%
  mutate(
    Mean.Discharge.clean = zoo::na.approx(mean_discharge, na.rm = FALSE)
  )

# Calculate the 7-day moving average, filling any remaining NAs
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver.clean %>%
  arrange(dateTime) %>%
  mutate(
    Mean.Discharge.7DayAvg = rollapply(
      Mean.Discharge.clean,
      width = 7,
      FUN = mean,
      align = "right",
      fill = NA
    )
  ) %>%
  # Fill remaining NAs in the 7-day moving average
  mutate(Mean.Discharge.7DayAvg = zoo::na.approx(Mean.Discharge.7DayAvg, na.rm = FALSE))

# If any missing values exist, replace them with a default value or use interpolation
CleanedDataPigeonRiver.clean <- CleanedDataPigeonRiver.clean %>%
  mutate(
    Mean.Discharge.7DayAvg = ifelse(
      is.na(Mean.Discharge.7DayAvg),
      mean(Mean.Discharge.7DayAvg, na.rm = TRUE), # Replace NAs with column mean
      Mean.Discharge.7DayAvg
    )
  )

# Generate a daily time series object for the 7-day moving average
s_year <- year(first(CleanedDataPigeonRiver.clean$dateTime))
s_month <- month(first(CleanedDataPigeonRiver.clean$dateTime))
f_year <- year(last(CleanedDataPigeonRiver.clean$dateTime))
f_month <- month(last(CleanedDataPigeonRiver.clean$dateTime))

# Re-create the daily time series object
CleanedDataPigeonRiver.daily.7DayAvg.ts <- ts(
  CleanedDataPigeonRiver.clean$Mean.Discharge.7DayAvg,
  start = c(s_year, s_month),
  end = c(f_year, f_month),
  frequency = 365
)

# Verify no missing values in the time series object
sum(is.na(CleanedDataPigeonRiver.daily.7DayAvg.ts))

# Decompose the 7-day moving average time series
CleanedDataPigeonRiver.daily.7DayAvg.decomp <- stl(
  CleanedDataPigeonRiver.daily.7DayAvg.ts,
  s.window = "periodic"
)

# Plot the components of the 7-day moving average decomposition
plot(
  CleanedDataPigeonRiver.daily.7DayAvg.decomp,
  main = "Decomposition of 7-Day Moving Average of Daily Discharge for Pigeon River"
)


```


These graphs are interesting, and some of them show 2024 as an outlier but they
don't say by exactly how much Helen was an outlier.

We want to be able to say with statistical certainty how unusual Helene was as
an event.

We will use the Hazen Method to find the recurrence intervals for Helene within
these parameters. We will do this below.


First we find the recurrence intervals of 7 days precipitation events. We know from
our first graphs that assuming Helen is our 7 day averaged maximum for 2024
is a reasonable assumption:

<span style="font-size:18px; font-weight:bold;">Helene Recurrence Intervals</span>

**Sugar Grove Hazen Precipitation**

```{r Sugar Grove Precip Hazen, warning=FALSE}
SugarGrovePrecip <- CleanedDataSugarGrove %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

SugarGrovePrecip <- SugarGrovePrecip %>%
  mutate(Year = year(dateTime))

SugarGrovePrecip <- SugarGrovePrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))

SugarGrovePrecip <- SugarGrovePrecip %>% 
  filter(MaxEvent > 0)

SugarGrovePrecip <- SugarGrovePrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

SugarGrovePrecipHazen2024 <- SugarGrovePrecip %>% filter(Year == 2024)
print(SugarGrovePrecipHazen2024)


```


**Asheville Hazen Precipitation**
```{r Asheville Precipitation Hazen, warning=FALSE}
AshevillePrecip <- CleanedDataAsheville %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

AshevillePrecip <- AshevillePrecip %>%
  mutate(Year = year(dateTime))

AshevillePrecip <- AshevillePrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm = TRUE))

AshevillePrecip <- AshevillePrecip %>% 
  filter(MaxEvent > 0)

AshevillePrecip <- AshevillePrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100 * (2 * rank - 1)) / (2 * n()))) %>% 
  mutate(hazen = 100 / Fa_Value)

AshevillePrecipHazen2024 <- AshevillePrecip %>% filter(Year == 2024)
print(AshevillePrecipHazen2024)


```

The Hazen number tells us that the 7 day average discharge observed in
Asheville in 2024 was a 54 year event, in other words it can be expected to
happen once every 54 years, or about a 2% chance of happening any one year. That
gives us a statistical number to match to our observed outliers. Let's find the 
recurrence intervals for 2024 for our other sites

**Pigeon River Hazen Precipitation**

```{r Pigeon River Precip Hazen, warning=FALSE}
PigeonRiverPrecip <- CleanedDataPigeonRiver %>% 
  mutate(moving_avg = rollapply(`Total Daily Precipitation (Inches)`, width = 7, FUN = mean, align = "center", fill = NA)) 

PigeonRiverPrecip <- PigeonRiverPrecip %>%
  mutate(Year = year(dateTime))

PigeonRiverPrecip <- PigeonRiverPrecip %>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm = TRUE))

PigeonRiverPrecip <- PigeonRiverPrecip  %>% 
  filter(MaxEvent > 0)

PigeonRiverPrecip <- PigeonRiverPrecip %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100 * (2 * rank - 1)) / (2 * n()))) %>% 
  mutate(hazen = 100 / Fa_Value)

PigeonRiverPrecipHazen2024 <- PigeonRiverPrecip %>% filter(Year == 2024)

print(PigeonRiverPrecipHazen2024)
```

In Pigeon River, it Helene was a 52 year event in terms of total precipitation,
but in Sugar Grove only an 18 year event, meaning Helene's
precipitation was less unusual in Sugar Grove than in Asheville or Pigeon River.


Let's now automate this function for our other two parameters and run our sites.

**Automated Discharge**

```{r}

Hazen2024Discharge.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
CleanDataFlood<-CleanData%>% 
  mutate(moving_avg = rollapply(`Mean Discharge`, width = 7, FUN = mean, align = "center", fill = NA)) 

CleanDataFlood<- CleanDataFlood%>%
  mutate(Year = year(dateTime))

CleanDataFlood <- CleanDataFlood%>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))

CleanDataFlood <- CleanDataFlood %>% 
  filter(MaxEvent > 0)

CleanDataFlood <- CleanDataFlood %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

CleanDataFlood2024 <- CleanDataFlood %>% filter(Year == 2024)


return(CleanDataFlood2024)
}

```

*And now we run our sites through our discharge function*

```{r}

SugarGroveDischargeHazen2024 <- Hazen2024Discharge.Function(siteNumberSugarGrove, pcode, scode)

AshevilleDischargeHazen2024 <- Hazen2024Discharge.Function(siteNumberAsheville, pcode, scode)

PigeonRiverDischargeHazen2024 <- Hazen2024Discharge.Function(siteNoPigeonRiver, pcode, scode)

SugarGroveDischargeHazen2024
AshevilleDischargeHazen2024
PigeonRiverDischargeHazen2024



```

*Now our function for Gage Height*

```{r}
Hazen2024GageHeight.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
CleanDataGageHeight<-CleanData%>% 
  mutate(moving_avg = rollapply(`Mean Gage Height`, width = 7, FUN = mean, align = "center", fill = NA)) 

CleanDataGageHeight<- CleanDataGageHeight%>%
  mutate(Year = year(dateTime))

CleanDataGageHeight <- suppressWarnings(CleanDataGageHeight%>% 
  group_by(Year) %>%                  
  summarize(MaxEvent = max(moving_avg, na.rm=T))
)

CleanDataGageHeight <- CleanDataGageHeight %>% 
  filter(MaxEvent > 0)

CleanDataGageHeight <- CleanDataGageHeight %>% 
  arrange(desc(MaxEvent)) %>% 
  mutate(rank = row_number()) %>% 
  mutate(Fa_Value = ((100*(2*rank - 1))/(2*n()))) %>% 
  mutate(hazen = 100/Fa_Value)

CleanDataGageHeight2024 <- CleanDataGageHeight %>% filter(Year == 2024)


return(CleanDataGageHeight2024)
}

#testing function

SugarGroveGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNumberSugarGrove, pcode, scode)

AshevilleGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNumberAsheville, pcode, scode)

PigeonRiverGageHeightHazen2024<-
  Hazen2024GageHeight.Function(siteNoPigeonRiver, pcode, scode)

SugarGroveGageHeightHazen2024
AshevilleGageHeightHazen2024
PigeonRiverGageHeightHazen2024

```



**Map of Gage Sites**

I would like to be able to now see geographically how unusual Helene was for each
site:
```{r}



library(RColorBrewer)

# Create Data Frame with Hazen Values
hazen_data <- data.frame(
  Location = c("Pigeon River", "Asheville", "Sugar Grove"),
  Latitude = c(35.554, 35.595, 36.267),
  Longitude = c(-82.982, -82.551, -81.785),
  Precipitation_Hazen = sapply(list(PigeonRiverPrecipHazen2024$hazen, 
                                    AshevillePrecipHazen2024$hazen, 
                                    SugarGrovePrecipHazen2024$hazen), function(x) as.numeric(unlist(x))),
  Discharge_Hazen = sapply(list(PigeonRiverDischargeHazen2024$hazen, 
                                 AshevilleDischargeHazen2024$hazen, 
                                 SugarGroveDischargeHazen2024$hazen), function(x) as.numeric(unlist(x))),
  GageHeight_Hazen = sapply(list(PigeonRiverGageHeightHazen2024$hazen, 
                                  AshevilleGageHeightHazen2024$hazen, 
                                  SugarGroveGageHeightHazen2024$hazen), function(x) as.numeric(unlist(x)))
)





color_pal <- colorNumeric(
  palette = c("blue", "green", "red"),  # Custom palette: low -> medium -> high
  domain = range(
    hazen_data$Precipitation_Hazen,
    hazen_data$Discharge_Hazen,
    hazen_data$GageHeight_Hazen,
    na.rm = TRUE
  ),
  reverse = FALSE
)


leaflet(hazen_data) %>%
  addTiles() %>%
  
  # Precipitation Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = ~color_pal(Precipitation_Hazen),
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Precipitation Hazen: ", round(Precipitation_Hazen, 2)
    ),
    radius = 10, group = "Precipitation"
  ) %>%
  
  # Discharge Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = ~color_pal(Discharge_Hazen),
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Discharge Hazen: ", round(Discharge_Hazen, 2)
    ),
    radius = 10, group = "Discharge"
  ) %>%
  
  # Gage Height Layer
  addCircleMarkers(
    ~Longitude, ~Latitude,
    color = ~color_pal(GageHeight_Hazen),
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Gage Height Hazen: ", round(GageHeight_Hazen, 2)
    ),
    radius = 10, group = "Gage Height"
  ) %>%
  
  # Add Layer Controls
  addLayersControl(
    overlayGroups = c("Precipitation", "Discharge", "Gage Height"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  
  # Add a Legend
  addLegend(
    "bottomright",
    pal = color_pal,
    values = range(
      hazen_data$Precipitation_Hazen,
      hazen_data$Discharge_Hazen,
      hazen_data$GageHeight_Hazen,
      na.rm = TRUE
    ),
    title = "Hazen Value (Recurrence Interval)",
    opacity = 0.7
  )


```

This is cool! I want to add a lot more sites now that I have this functioning.
I go and download more sites in Western NC, by randomly selecting one per county
for every counry in Western NC and find the Hazen numbers for each parameter
```{r}




```



I then add them to the heatmap
```{r}

#install.packages("leaflet.extras")
library(leaflet.extras)

leaflet(hazen_data) %>%
  addTiles() %>%
  addHeatmap(
    lng = ~Longitude,
    lat = ~Latitude,
    intensity = ~Precipitation_Hazen,
    blur = 20,
    max = max(hazen_data$Precipitation_Hazen, na.rm = TRUE),
    radius = 15
  ) %>%
  addLegend(
    "bottomright",
    pal = color_pal,
    values = hazen_data$Precipitation_Hazen,
    title = "Precipitation Hazen",
    opacity = 0.7
  )


leaflet(hazen_data) %>%
  addTiles() %>%
  addCircleMarkers(
    ~Longitude, ~Latitude,
    fillColor = ~color_pal(Discharge_Hazen),
    fillOpacity = 0.8,
    stroke = FALSE,
    radius = 10,
    label = ~paste0(
      "Location: ", Location, "<br>",
      "Discharge Hazen: ", round(Discharge_Hazen, 2)
    )
  ) %>%
  addLegend(
    "bottomright",
    pal = color_pal,
    values = hazen_data$Discharge_Hazen,
    title = "Discharge Hazen",
    opacity = 0.7
  )



```