---
title: "Scrape.it for Final Project"
author: "Sam Tolbert"
date: "2024-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rvest)
library(purrr)
library(lubridate)
library(here)
library(dplyr)

install.packages("dataRetrieval")
library(dataRetrieval)



```

Retrieving data for Sugar Grove Gage from USGS using USGS's dataRetrieval package.
I find its siteNumber is 03479000 from "https://waterdata.usgs.gov/monitoring-location/03479000/#parameterCode=
00060&period=P7D&showMedian=false".
```{r}

siteNumberSugarGrove<-('03479000')

#I find the MetaData for the site
SugarGroveMeta <- whatNWISsites(siteNumbers = siteNumberSugarGrove)
print(SugarGroveMeta) 

#I then use sitedata function to look at what data is available
SugarGroveSiteData<-suppressWarnings(whatNWISdata(siteNumber=siteNumberSugarGrove))
```

Looking at what data is available, I see that the parameter codes (parmcodes)
for stream discharge (00060) , precipitation (00045), and gage height (0065).
I also see that our relevant stat codes are daily mean (0003) and sum (00006)
#

```{r warning= FALSE message=FALSE}
pcodes<-c('00060','00045','00065') #parameters discharge, precip, and gage height
scode<-c('00003', '00006') #daily mean and daily accumulated


SugarGroveData<-readNWISdata( #dataRetrival function from USGS API
  siteNumbers = siteNumberSugarGrove,
  parameterCd= pcodes,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)
  
View(SugarGroveData)  
```

This is my data! However, I need to clean this up. 
First I want to remove the 'cd' columns that indicate if the data is official or 
provisional as this won't have a significant effect on our findings. 
We can mention that in our discussion of the meta data but its not 
worth a column in our df. 

Then I want to clearly label each column with its parameter and stat code:

```{r}
CleanedDataSugarGrove<-SugarGroveData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
View(CleanSugarGroveData)


```

I want to repeat this data retrieval and cleaning for two additional sites,
French Broad River in Asheville in Southwestern NC (03451500) 
and East Fork Pigeon River in far Southwestern NC near Canton (03456500) for
examinations of 3 different rivers in different parts of the state.

First I check the metadata and sitedata to see if similar data and parameters
are availale to compare to Sugar Grove

```{r}

#First the Asheville site

siteNumberAsheville<-('03451500')

#I find the MetaData for the site
AshevilleMeta <- whatNWISsites(siteNumbers = siteNumberAsheville)
print(AshevilleMeta) 

#I then use sitedata function to look at what data is available
AshevilleSiteData<-whatNWISdata(siteNumber=siteNumberAsheville)

```

Then I decide I want to create an easy data retrieval function for my chosen
parameters based on my previous retreival.

```{r}

WaterData.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
df_name<-paste0('CleanData_', siteNo)
assign(df_name, CleanData, envir= .GlobalEnv)

return(invisible(CleanData))
}
```

I run this function for Asheville's siteNo
```{r}
siteNoAsheville<-'03451500'

siteNo<-siteNoAsheville
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataAsheville<-WaterData.Function(siteNo, pcode, scode)
View(CleanedDataAsheville)


```

I then do the same for the Pigeon River Site and its site number

```{r}

siteNoPigeonRiver<-'03456500'

siteNo<-siteNoPigeonRiver
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataPigeonRiver<-WaterData.Function(siteNo, pcode, scode)
View(CleanedDataPigeonRiver)
  

```

##Data Visualization## 
```{r}

library(ggplot2)

ggplot(CleanedDataAsheville, aes(
  x = dateTime, 
  y = `Mean Discharge`)) +
  geom_line(color = "green", size = 1) +  
  geom_point(color = "darkgreen", size = 2) +  
  labs(
    title = "Mean Daily Discharge Over Time",
    subtitle = "Asheville(1954-2024)",
    x = "Date",
    y = "Mean Discharge (cubic meters per second)"
  ) +
  theme_minimal() 

ggplot(CleanedDataPigeonRiver, aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="purple")+
  geom_point(color="maroon")+
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Pigeon River(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()


ggplot(CleanedDataSugarGrove, aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="blue")+
  geom_point(color="darkblue")+
  labs(
    title = "Mean Daily Discharge Over time",
    subtitle = "Sugar Grove(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)"
  )+ theme_minimal()
  


#note: what I am noticing is that we have a lot of daily data which is good when
#we take a closer look at Helene. I am going to do some work looking at the monthly
#mean daily precip for now- We will need to cut data in the future 1980-2024
#these do show the outliers which is kind of neat 


```

##Pigeon River Visualizations##
```{r}



PRMonthlyMean <- CleanedDataPigeonRiver %>%
  mutate(dateTime = floor_date(dateTime, "month")) %>% #first day of the month
  group_by(dateTime) %>%
  summarize(`Mean Discharge` = mean(`Mean Discharge`, na.rm = TRUE))

#What I did here was aggregate the data so now we have a monthly mean 
#I made it just show up as the first day of the month 
#getting the mean value for the specific month 
#basically I am just playing with less data 

ggplot(PRMonthlyMean, aes(
  x= dateTime,
  y= `Mean Discharge`)) +
  geom_line(color="purple")+
  geom_point(color="maroon")+
  labs(
    title = "Mean Monthly Discharge Over time",
    subtitle = "Pigeon River(1954-2024)",
    x= "Date",
    y= "Mean Discharge (cubic meters per second)")+ 
    theme_minimal()

str(CleanedDataPigeonRiver)

CleanedDataPigeonRiver <- CleanedDataPigeonRiver %>%
  mutate(dateTime = as.Date(dateTime))  

#okay I still don't like the visualization 
#I think we need to do a log scale 



```

##Pigeon River seasonal breakdown and log regression## 
```{r}
#I am now breaking the data into hurricane and frontal seasons 
#hurricane season is July to November 
#frontal season is December- June
#This info is from previous Hydrology project 

SeasonalCleanedDataPigeonRiver <- CleanedDataPigeonRiver %>%
  mutate(season = case_when(
    month(dateTime) %in% 7:11 ~ "Hurricane",  # July to November
    month(dateTime) %in% c(12, 1:6) ~ "Frontal"  # December to June
  ))

head(SeasonalCleanedDataPigeonRiver)

#Pigeon River Data is now in frontal and hurricane seasons with a new column
#called seasons 

PRHurricaneData <- SeasonalCleanedDataPigeonRiver %>%
  filter(season == "Hurricane")

PRFrontalData <- SeasonalCleanedDataPigeonRiver %>%
  filter(season == "Frontal")

#plotting the two seasons 

ggplot(SeasonalCleanedDataPigeonRiver, aes(
  x = dateTime, 
  y = `Mean Discharge`, 
  color = season)) +
  geom_point(size = 1) +  # Scatterplot
  labs(
    title = "Mean Daily Discharge During Hurricane and Frontal Seasons",
    subtitle = "Pigeon River (1954-2024",
    x = "Date",
    y = "Mean Daily Discharge (cfs)",
    color = "Season"
  ) +
  scale_color_manual(values = c("Hurricane" = "purple", "Frontal" = "blue"))  


ggplot(PRHurricaneData, aes(
  x = dateTime, 
  y = `Mean Discharge`)) +
  geom_point(color = "purple") +  # Points for yearly average precipitation
  geom_smooth(method = "lm", color = "black", se = TRUE) +  # Linear regression with confidence interval
  labs(
    title = "Trend in Precipitation During Hurricane Seasons",
    x = "Year",
    y = "Mean Precipitation (mm)"
  ) +
  theme_minimal()

#plotting a linear regression for hurricane season on the Pigeon River 

#log data 

ggplot(PRHurricaneData, aes(
  x = dateTime, y = log(`Mean Discharge`))) + 
  geom_point(color = "purple") +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  labs(
    title = "Log-Transformed Trend in Precipitation During Hurricane Seasons",
    subtitle = "Pigeon River (1954-2024)",
    x = "Year",
    y = "Log of Mean Precipitation (mm)"
  ) +
  theme_minimal()









```




