---
title: "Scrape.it for Final Project"
author: "Sam Tolbert"
date: "2024-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rvest)
library(purrr)
library(lubridate)
library(here)
library(dplyr)

#install.packages("dataRetrieval")
library(dataRetrieval)



```

Retrieving data for Sugar Grove Gage from USGS using USGS's dataRetrieval package.
I find its siteNumber is 03479000 from "https://waterdata.usgs.gov/monitoring-location/03479000/#parameterCode=
00060&period=P7D&showMedian=false".
```{r}

siteNumberSugarGrove<-('03479000')

#I find the MetaData for the site
SugarGroveMeta <- whatNWISsites(siteNumbers = siteNumberSugarGrove)
print(SugarGroveMeta) 

#I then use sitedata function to look at what data is available
SugarGroveSiteData<-suppressWarnings(whatNWISdata(siteNumber=siteNumberSugarGrove))
```

Looking at what data is available, I see that the parameter codes (parmcodes)
for stream discharge (00060) , precipitation (00045), and gage height (0065).
I also see that our relevant stat codes are daily mean (0003) and sum (00006)
#

```{r warning= FALSE message=FALSE}
pcodes<-c('00060','00045','00065') #parameters discharge, precip, and gage height
scode<-c('00003', '00006') #daily mean and daily accumulated


SugarGroveData<-readNWISdata( #dataRetrival function from USGS API
  siteNumbers = siteNumberSugarGrove,
  parameterCd= pcodes,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)
  
View(SugarGroveData)  
```

This is my data! However, I need to clean this up. 
First I want to remove the 'cd' columns that indicate if the data is official or 
provisional as this won't have a significant effect on our findings. 
We can mention that in our discussion of the meta data but its not 
worth a column in our df. 

Then I want to clearly label each column with its parameter and stat code:

```{r}
CleanedDataSugarGrove<-SugarGroveData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
View(CleanSugarGroveData)


```

I want to repeat this data retrieval and cleaning for two additional sites,
French Broad River in Asheville in Southwestern NC (03451500) 
and East Fork Pigeon River in far Southwestern NC near Canton (03456500) for
examinations of 3 different rivers in different parts of the state.

First I check the metadata and sitedata to see if similar data and parameters
are availale to compare to Sugar Grove

```{r}

#First the Asheville site

siteNumberAsheville<-('03451500')

#I find the MetaData for the site
AshevilleMeta <- whatNWISsites(siteNumbers = siteNumberAsheville)
print(AshevilleMeta) 

#I then use sitedata function to look at what data is available
AshevilleSiteData<-whatNWISdata(siteNumber=siteNumberAsheville)

```

Then I decide I want to create an easy data retrieval function for my chosen
parameters based on my previous retreival.

```{r}

WaterData.Function<-function(siteNo, pcode, scode){

SiteData<-readNWISdata(
  siteNumbers = siteNo,
  parameterCd= pcode,
  statCd= scode,
  startDate= ymd('1900-01-01'),
  endDate= Sys.Date()
)

  CleanData<-SiteData%>% 
select(-ends_with("cd")) %>% #removing cd columns that indiciate provisional or offical data
  rename(
  'Total Daily Precipitation (Inches)' = X_00045_00006, #renaming columns
    'Mean Discharge' = X_00060_00003,
    'Mean Gage Height' = X_00065_00003
  )
  
df_name<-paste0('CleanData_', siteNo)
assign(df_name, CleanData, envir= .GlobalEnv)

return(invisible(CleanData))
}
```

I run this function for Asheville's siteNo
```{r}
siteNoAsheville<-'03451500'

siteNo<-siteNoAsheville
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataAsheville<-WaterData.Function(siteNo, pcode, scode)
View(CleanedDataAsheville)


```

I then do the same for the Pigeon River Site and its site number

```{r}

siteNoPigeonRiver<-'03456500'

siteNo<-siteNoPigeonRiver
pcode<-c('00060','00045','00065')
scode<-c('00003', '00006')

CleanedDataPigeonRiver<-WaterData.Function(siteNo, pcode, scode)
View(CleanedDataPigeonRiver)
  

```